---
title: "LME_Oregon_Cog_and_symptoms_AIO"
author: "Jack Scott"
date: "2026-07-22"
output:
  html_document:
    df_print: paged
---

# Setup
```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir ='/media/hcs-sci-psy-narun/Jack/First_level_model_aligned/Stacked_model_ridge_v4_EXTREME_conserve_a2000_keeper')
``` 

# Packages
```{r packages}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(r2mlm)
library(glmm.hp)
library(broom.mixed)
```

# Data Preparation
```{r data_prep}
# Load Neuroimaging/Cognition Data
cognition_data <- readr::read_csv("/media/hcs-sci-psy-narun/Jack/First_level_model_aligned/Stacked_model_ridge_v4_EXTREME_conserve_a2000_keeper/stacked_ridge_predictions.csv")

# Load Phenotype Data (ADHD Diagnosis)
phenotype_data <- read_csv("/media/hcs-sci-psy-narun/Jack/HMM_ELM_data_prep/phase4_statistical_analysis/analysis_ready_dataset.csv")

# Load Inattention/Hyperactivity Matched Datasets
inattention_scores <- readr::read_csv("/media/hcs-sci-psy-narun/Jack/LME_cog_inattention/inattention_scores_matched.csv")
hyperactivity_scores <- readr::read_csv("/media/hcs-sci-psy-narun/Jack/LME_cog_hyperactivity/hyperactivity_scores_matched.csv")

# Pre-processing & Merging
analysis_data <- cognition_data %>%
  # Extract metadata
  mutate(participant_id = as.character(str_extract(subject_id, "^[^_]+")),
         Age = as.numeric(str_extract(subject_id, "(?<=_Age)\\d+"))) %>%
  # Merge Phenotypes (primary: timepoint-specific join)
  left_join(
    phenotype_data %>%
      mutate(subject_id_match = paste0(subject_id, "_Age", sprintf("%03d", as.integer(age_months)))) %>%
      select(subject_id_match, final_phenotype, sex),
    by = c("subject_id" = "subject_id_match")
  ) %>%
  # Merge Phenotypes (fallback: participant-level join for age-mismatched timepoints)
  left_join(
    phenotype_data %>%
      distinct(subject_id, .keep_all = TRUE) %>%
      select(subject_id, final_phenotype_fb = final_phenotype, sex_fb = sex),
    by = c("participant_id" = "subject_id")
  ) %>%
  mutate(
    final_phenotype = coalesce(final_phenotype, final_phenotype_fb),
    sex = coalesce(sex, sex_fb)
  ) %>%
  select(-final_phenotype_fb, -sex_fb) %>%
  # Clean Diagnosis Label
  mutate(ADHDdiagnosis = factor(case_when(
    final_phenotype == 3 ~ "ADHD",
    final_phenotype == 1 ~ "Control",
    TRUE ~ NA_character_
  ), levels = c("Control", "ADHD"))) %>%
  
  # Merge Symptoms
  mutate(
    InattentionRaw = inattention_scores$inattention[match(subject_id, inattention_scores$subject_id)],
    HyperactivityRaw = hyperactivity_scores$hyperactivity[match(subject_id, hyperactivity_scores$subject_id)]
  ) %>%
  # Standardizing
  mutate(
    participant_id = as_factor(participant_id),
    Age0 = Age - min(Age, na.rm = TRUE),
    true_g_z = as.double(scale(true_g)),
    stacked_pred_z = as.double(scale(stacked_pred)),
    inattention_z = as.double(scale(InattentionRaw)),
    hyperactivity_z = as.double(scale(HyperactivityRaw)),
    sex_num = if_else(sex == "M", 0, 1)
  ) %>%
  # Within/Between Decomposition
  group_by(participant_id) %>%
    mutate(
      true_g_z_mean = mean(true_g_z, na.rm = TRUE),
      stacked_pred_z_mean = mean(stacked_pred_z, na.rm = TRUE),
      InattentionMean = mean(inattention_z, na.rm = TRUE),
      HyperactivityMean = mean(hyperactivity_z, na.rm = TRUE),
      age_mean = mean(Age0, na.rm = TRUE),
      count_timepoints = n()
    ) %>%
  ungroup() %>%
  mutate(
    true_g_z_dev = true_g_z - true_g_z_mean,
    stacked_pred_z_dev = stacked_pred_z - stacked_pred_z_mean,
    InattentionDeviation = inattention_z - InattentionMean,
    HyperactivityDeviation = hyperactivity_z - HyperactivityMean,
    age_deviation = Age0 - age_mean
  ) %>%
  # NOTE: filter applied separately below to preserve full dataset for sensitivity analyses
  identity()

# Full dataset (all participants and timepoints) retained for sensitivity analyses
analysis_data_full <- analysis_data

# Longitudinal subset for main LME analyses
analysis_data <- analysis_data_full %>%
  filter(count_timepoints >= 2)

cat("Full dataset:          ", nrow(analysis_data_full), "observations from", n_distinct(analysis_data_full$participant_id), "participants.\n")
cat(mean(analysis_data_full$Age), "mean age of participants", sd(analysis_data_full$Age), "standard deviation\n")
adhd_data <- analysis_data_full %>%
  filter(final_phenotype == 3) %>%
  distinct(participant_id, .keep_all = TRUE)

cat(sprintf("ADHD subsample: N=%d\n", nrow(adhd_data)))
cat(sprintf("Mean age: %.1f months (SD=%.1f, range %d-%d)\n",
            mean(adhd_data$Age, na.rm = TRUE),
            sd(adhd_data$Age, na.rm = TRUE),
            min(adhd_data$Age, na.rm = TRUE),
            max(adhd_data$Age, na.rm = TRUE)))

sex_adhd <- adhd_data %>% count(sex)
cat("Sex breakdown:\n")
for (i in seq_len(nrow(sex_adhd))) {
  cat(sprintf("  %s: %d (%.1f%%)\n",
              sex_adhd$sex[i],
              sex_adhd$n[i],
              sex_adhd$n[i] / nrow(adhd_data) * 100))
}
cat("Longitudinal subset:   ", nrow(analysis_data), "observations from", n_distinct(analysis_data$participant_id), "participants.\n")



```
```{r double check counts}
analysis_data_full %>%
  distinct(participant_id, .keep_all = TRUE) %>%
  count(final_phenotype) %>%
  print()

cat("Total:", n_distinct(analysis_data_full$participant_id), "\n")

# ---- Longitudinal sample: sex and diagnosis breakdown (Reviewer 4, Q4) ----
# Both participant-level and timepoint-level counts are reported, so that a
# potentially sparser ADHD contribution at the timepoint level is visible
# alongside the participant headcount. Percentages use the full N as the
# denominator (missing values included), matching the existing methods sentence.

n_ppt <- n_distinct(analysis_data$participant_id)
n_tp  <- nrow(analysis_data)

# One row per participant. Sex is time-invariant, so the first non-missing value
# is taken. The phenotype is collapsed with a stable rule (assigned ADHD if ever
# classified ADHD across visits); see the caveat below.
ppt_level <- analysis_data %>%
  group_by(participant_id) %>%
  summarise(
    sex   = na.omit(sex)[1],
    pheno = if (any(final_phenotype == 3, na.rm = TRUE)) 3L else na.omit(final_phenotype)[1],
    .groups = "drop"
  )

cat(sprintf("\nLongitudinal sample: %d participants, %d timepoints\n\n", n_ppt, n_tp))

cat("Sex, participant level:\n")
print(ppt_level %>% count(sex) %>% mutate(pct = round(100 * n / n_ppt, 2)))

cat("\nDiagnosis (final_phenotype), participant level [stable rule]:\n")
print(ppt_level %>% count(pheno) %>% mutate(pct = round(100 * n / n_ppt, 2)))

cat("\nSex, timepoint level:\n")
print(analysis_data %>% count(sex) %>% mutate(pct = round(100 * n / n_tp, 2)))

cat("\nDiagnosis (final_phenotype), timepoint level [per visit]:\n")
print(analysis_data %>% count(final_phenotype) %>% mutate(pct = round(100 * n / n_tp, 2)))
```

```{r longitudinal descriptives}
# Participant-level breakdown for longitudinal subsample (2+ timepoints)
longi_pts <- analysis_data %>%
  group_by(participant_id) %>%
  filter(count_timepoints >= 2) %>%
  distinct(participant_id, .keep_all = TRUE) %>%
  ungroup()

cat(sprintf("Total participants (2+ timepoints): %d\n", nrow(longi_pts)))

sex_counts <- longi_pts %>% count(sex)
cat("Sex breakdown (participant level):\n")
for (i in seq_len(nrow(sex_counts))) {
  cat(sprintf("  %s: %d (%.1f%%)\n",
              sex_counts$sex[i],
              sex_counts$n[i],
              sex_counts$n[i] / nrow(longi_pts) * 100))
}

pheno_counts <- longi_pts %>% count(final_phenotype)
cat("Phenotype breakdown (participant level):\n")
for (i in seq_len(nrow(pheno_counts))) {
  cat(sprintf("  %s: %d (%.1f%%)\n",
              pheno_counts$final_phenotype[i],
              pheno_counts$n[i],
              pheno_counts$n[i] / nrow(longi_pts) * 100))
}
```


```{r}
# Calculate Pearson correlation between deviation scores
pearson_correlation_deviations <- cor(analysis_data$true_g_z_dev, 
                          analysis_data$stacked_pred_z_dev, 
                          method = "pearson",
                          use = "complete.obs")

# Get unique participant-level data (since means are repeated for each observation)
participant_means <- analysis_data[!duplicated(analysis_data$participant_id), ]

# Create a scatterplot to visualize the relationship
library(ggplot2)
ggplot(analysis_data, aes(x = true_g_z_dev, y = stacked_pred_z_dev)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "True g (deviation scores)", 
       y = "Predicted g (deviation scores)",
       title = "Within-Person Relationship: Observed vs Predicted g",
       subtitle = paste("Pearson r =", round(pearson_correlation_deviations, 3))) +
  theme_minimal()

# Calculate Pearson correlation between participant means
pearson_means <- cor(participant_means$true_g_z_mean, 
                    participant_means$stacked_pred_z_mean, 
                    method = "pearson",
                    use = "complete.obs")

# Create a scatterplot for participant means
ggplot(participant_means, aes(x = true_g_z_mean, y = stacked_pred_z_mean)) +
  geom_point(alpha = 0.7, size = 2, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red", fill = "red", alpha = 0.2) +  
  # Add correlation statistics in a properly sized box (lower right)
  annotate("label", 
           x = max(participant_means$true_g_z_mean, na.rm = TRUE) * 0.95, 
           y = min(participant_means$stacked_pred_z_mean, na.rm = TRUE) * 0.95,
           label = paste("Pearson r =", round(pearson_means, 3), 
                         "\nn =", nrow(participant_means)),
           size = 3,
           fill = "white", 
           color = "black",
           alpha = 0.9,
           label.size = 0.15) +  
  labs(x = "Observed g (participant means)", 
       y = "Predicted g (participant means)",
       title = "Between-Person Relationship: Observed vs Predicted g") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        # Note: Changed size to linewidth to avoid ggplot2 deprecation warnings
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5))
``` 


# Model 1: ADHD Additive
```{r m1_adhd_additive}
m1_adhd_add <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + ADHDdiagnosis + (1 | participant_id), 
                    data = analysis_data, REML = FALSE)
summary(m1_adhd_add)
```

# Model 2: ADHD Interactive
```{r m2_adhd_interactive}
m2_adhd_int <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + ADHDdiagnosis + 
                    stacked_pred_z_dev:ADHDdiagnosis + stacked_pred_z_mean:ADHDdiagnosis + (1 | participant_id), 
                    data = analysis_data, REML = FALSE)
summary(m2_adhd_int)
```

# Model 3: Neuroimaging-only
```{r m3_neuroimaging}
m3_neuro <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m3_neuro)
```




# Model 4: Age-only
```{r m4_age}
m4_age <- lmer(true_g_z ~ age_deviation + age_mean + (1 | participant_id), 
               data = analysis_data, REML = FALSE)
summary(m4_age)
```

# Model 5: Neuroimaging-and-age
```{r m5_neuro_age}
m5_neuro_age <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + age_deviation + age_mean + (1 | participant_id), 
                     data = analysis_data, REML = FALSE)
summary(m5_neuro_age)

print("\n ===DEBUG===")
# Diagnostic: Print full r2mlm structure
# r2_full <- r2mlm::r2mlm(m5_neuro_age)
# print("Full r2mlm output:")
# print(r2_full)
# print("\nNames in r2mlm object:")
# print(names(r2_full))
# print("\nStructure:")
# str(r2_full)

```

# Model 6: Hyperactivity
```{r m6_hyperactivity}
m6_hyper <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + HyperactivityDeviation + HyperactivityMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m6_hyper)
```
# hyperactivity only
```{r m6p5_hyperactivity_only}
m6p5_hyper <- lmer(true_g_z ~ HyperactivityDeviation + HyperactivityMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m6p5_hyper)
```


# Model 7: Inattention
```{r m7_inattention}
m7_inatt <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + InattentionDeviation + InattentionMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m7_inatt)
```

```{r m7p5_inattention}
m7p5_inatt <- lmer(true_g_z ~ InattentionDeviation + InattentionMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m7p5_inatt)
```

# ADHD Model Comparison
```{r adhd_comparison}
# Formal comparison
comp_adhd <- anova(m1_adhd_add, m2_adhd_int)
print(comp_adhd)

# BIC Difference logic
delta_bic <- BIC(m2_adhd_int) - BIC(m1_adhd_add)
cat("\nDelta BIC (Interactive - Additive):", round(delta_bic, 2), "\n")
if(delta_bic > 0) {
  cat("Evidence favors the simpler ADDITIVE model.\n")
} else {
  cat("Evidence favors the INTERACTIVE model.\n")
}
```

# Core Models Comparison
```{r core_comparison}
comp_biomarkers_age <- anova(m3_neuro, m4_age, m5_neuro_age)
print(comp_biomarkers_age)
```

# SJPlot Tables: Core Models
```{r sjplot_core}
tab_model(m3_neuro, m4_age, m5_neuro_age, 
          dv.labels = c("Neuroimaging Only", "Age Only", "Neuroimaging-and-Age"),
          pred.labels = c(
            "(Intercept)" = "Intercept",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "age_deviation" = "age<sub>deviation</sub>",
            "age_mean" = "age<sub>mean</sub>"
          ),
          show.aic = TRUE,
          show.dev = TRUE,  # Shows deviance instead of BIC
          title = "Model Comparison: Neuroimaging and Age",
          file = "core_models")
```

# SJPlot Tables: ADHD Models
```{r sjplot_adhd}
tab_model(m1_adhd_add, m2_adhd_int, 
          dv.labels = c("ADHD Additive", "ADHD Interactive"),
          pred.labels = c(
            "(Intercept)" = "(Intercept)",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD" = "ADHD Diagnosis",
            "stacked_pred_z_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD"),
          show.aic = TRUE,
          show.dev = TRUE,
          title = "ADHD Diagnosis Comparison",
          file = "adhd_diagnosis")
```

# SJPlot Tables: Symptom models
```{r sjplot_hyp}
tab_model(m6p5_hyper, m6_hyper, 
          dv.labels = c("Hyperactivity only", "Neuroimaging-and-Hyperactivity"),
          pred.labels = c(
            "(Intercept)" = "Intercept",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD" = "ADHD Diagnosis",
            "stacked_pred_z_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD",
            "HyperactivityMean" = "Hyperactivity<sub>mean</sub>",
            "HyperactivityDeviation" = "Hyperactivity<sub>deviation</sub>"
          ),
          show.aic = TRUE,
          show.dev = TRUE,
          title = "Symptoms Comparison",
          file = "symptom_models")
```

```{r sjplot_inatt}
tab_model(m7p5_inatt, m7_inatt, 
          dv.labels = c("Inattention Only", "Neuroimaging-and-Inattention"),
          pred.labels = c(
            "(Intercept)" = "Intercept",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD" = "ADHD Diagnosis",
            "stacked_pred_z_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD",
            "InattentionMean" = "Inattention<sub>mean</sub>",
            "InattentionDeviation" = "Inattention<sub>deviation</sub>" 
          ),
          show.aic = TRUE,
          show.dev = TRUE,
          title = "Symptoms Comparison",
          file = "symptom_models")
```

# R2 Decomposition: Model 5
```{r m5_r2_calculation}
# Calculate R-squared for Model 5
r2_m5 <- r2mlm::r2mlm(m5_neuro_age)
print("Model 5 R-squared:")
print(r2_m5)
```

```{r m3_r2_calculation}
# Calculate R-squared for Model 3
r2_m3 <- r2mlm::r2mlm(m3_neuro)
print("Model 3 R-squared:")
print(r2_m3)
```


# Commonality Analysis: Model 5
```{r m5_commonality}
# Refit using lme4 for glmm.hp compatibility
m5_lme4 <- lme4::lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + age_deviation + age_mean + (1 | participant_id), 
                      data = analysis_data, REML = FALSE)

# Run commonality analysis
m5_commonality_results <- glmm.hp::glmm.hp(m5_lme4, type = "adjR2", commonality = TRUE)
print("Model 5 Commonality Results:")
print(m5_commonality_results)
```

# Adjust Commonality: Model 5
```{r m5_adjust_commonality}
# Extract commonality analysis results
m5_commonality_raw <- m5_commonality_results$commonality.analysis

# Remove total rows
is_total_row <- grepl("total|Total|TOTAL", rownames(m5_commonality_raw), ignore.case = TRUE)
m5_components_only <- m5_commonality_raw[!is_total_row, , drop = FALSE]

# Set negative values to zero
m5_components_positive <- ifelse(m5_components_only[,1] < 0, 0, m5_components_only[,1])
names(m5_components_positive) <- rownames(m5_components_only)

# Re-normalize to 100%
m5_sum_positive <- sum(m5_components_positive)
m5_components_adjusted_pct <- (m5_components_positive / m5_sum_positive) * 100

# Create adjusted results table
m5_commonality_adjusted <- data.frame(
  Component = rownames(m5_components_only),
  Original_Fraction = m5_components_only[,1],
  Original_Percent = m5_components_only[,2],
  Adjusted_Fraction = m5_components_positive,
  Adjusted_Percent = m5_components_adjusted_pct,
  stringsAsFactors = FALSE
)

# Add adjusted total row
m5_total_row <- data.frame(
  Component = "Total",
  Original_Fraction = sum(m5_components_only[,1]),
  Original_Percent = 100.00,
  Adjusted_Fraction = m5_sum_positive,
  Adjusted_Percent = 100.00,
  stringsAsFactors = FALSE
)

# Combine
m5_commonality_adjusted_final <- rbind(m5_commonality_adjusted, m5_total_row)
print("Model 5 Adjusted Commonality:")
print(m5_commonality_adjusted_final)
```

# Commonality Analysis: Model 6 (Hyperactivity)
```{r m6_commonality}
# Refit using lme4
m6_lme4 <- lme4::lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + HyperactivityDeviation + HyperactivityMean + (1 | participant_id), 
                      data = analysis_data, REML = FALSE)

# Run commonality analysis
m6_commonality_results <- glmm.hp::glmm.hp(m6_lme4, type = "adjR2", commonality = TRUE)
print("Model 6 Commonality Results:")
print(m6_commonality_results)
```

# Adjust Commonality: Model 6
```{r m6_adjust_commonality}
# Extract commonality analysis results
m6_commonality_raw <- m6_commonality_results$commonality.analysis

# Remove total rows
is_total_row <- grepl("total|Total|TOTAL", rownames(m6_commonality_raw), ignore.case = TRUE)
m6_components_only <- m6_commonality_raw[!is_total_row, , drop = FALSE]

# Set negative values to zero
m6_components_positive <- ifelse(m6_components_only[,1] < 0, 0, m6_components_only[,1])
names(m6_components_positive) <- rownames(m6_components_only)

# Re-normalize to 100%
m6_sum_positive <- sum(m6_components_positive)
m6_components_adjusted_pct <- (m6_components_positive / m6_sum_positive) * 100

# Create adjusted results table
m6_commonality_adjusted <- data.frame(
  Component = rownames(m6_components_only),
  Original_Fraction = m6_components_only[,1],
  Original_Percent = m6_components_only[,2],
  Adjusted_Fraction = m6_components_positive,
  Adjusted_Percent = m6_components_adjusted_pct,
  stringsAsFactors = FALSE
)

# Add adjusted total row
m6_total_row <- data.frame(
  Component = "Total",
  Original_Fraction = sum(m6_components_only[,1]),
  Original_Percent = 100.00,
  Adjusted_Fraction = m6_sum_positive,
  Adjusted_Percent = 100.00,
  stringsAsFactors = FALSE
)

# Combine
m6_commonality_adjusted_final <- rbind(m6_commonality_adjusted, m6_total_row)
print("Model 6 Adjusted Commonality:")
print(m6_commonality_adjusted_final)
```

# Commonality Analysis: Model 7 (Inattention)
```{r m7_commonality}
# Refit using lme4
m7_lme4 <- lme4::lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + InattentionDeviation + InattentionMean + (1 | participant_id), 
                      data = analysis_data, REML = FALSE)

# Run commonality analysis
m7_commonality_results <- glmm.hp::glmm.hp(m7_lme4, type = "adjR2", commonality = TRUE)
print("Model 7 Commonality Results:")
print(m7_commonality_results)
```

# Adjust Commonality: Model 7
```{r m7_adjust_commonality}
# Extract commonality analysis results
m7_commonality_raw <- m7_commonality_results$commonality.analysis

# Remove total rows
is_total_row <- grepl("total|Total|TOTAL", rownames(m7_commonality_raw), ignore.case = TRUE)
m7_components_only <- m7_commonality_raw[!is_total_row, , drop = FALSE]

# Set negative values to zero
m7_components_positive <- ifelse(m7_components_only[,1] < 0, 0, m7_components_only[,1])
names(m7_components_positive) <- rownames(m7_components_only)

# Re-normalize to 100%
m7_sum_positive <- sum(m7_components_positive)
m7_components_adjusted_pct <- (m7_components_positive / m7_sum_positive) * 100

# Create adjusted results table
m7_commonality_adjusted <- data.frame(
  Component = rownames(m7_components_only),
  Original_Fraction = m7_components_only[,1],
  Original_Percent = m7_components_only[,2],
  Adjusted_Fraction = m7_components_positive,
  Adjusted_Percent = m7_components_adjusted_pct,
  stringsAsFactors = FALSE
)

# Add adjusted total row
m7_total_row <- data.frame(
  Component = "Total",
  Original_Fraction = sum(m7_components_only[,1]),
  Original_Percent = 100.00,
  Adjusted_Fraction = m7_sum_positive,
  Adjusted_Percent = 100.00,
  stringsAsFactors = FALSE
)

# Combine
m7_commonality_adjusted_final <- rbind(m7_commonality_adjusted, m7_total_row)
print("Model 7 Adjusted Commonality:")
print(m7_commonality_adjusted_final)
```

# Export R2 Decomposition: Model 5
```{r export_m5_r2_decomposition}
# Export R2 Decomposition: Model 3 - CORRECTED
# Preserve the matrix structure (total/within/between columns)

r2_decomposition <- r2mlm::r2mlm(m3_neuro)

# Extract the Decompositions matrix
decomp_matrix <- r2_decomposition$Decompositions

# Create dataframe with proper row and column labels
decomp_df <- data.frame(
  component = rownames(decomp_matrix),
  total = decomp_matrix[, "total"],
  within = decomp_matrix[, "within"],
  between = decomp_matrix[, "between"],
  stringsAsFactors = FALSE
)

# Export for Python
write.csv(decomp_df, "m3_r2mlm_decomposition.csv", row.names = FALSE)
cat("Exported Model 3 R2 decomposition to m3_r2mlm_decomposition.csv\n")
```

# Export Adjusted Commonality: All Models
```{r export_commonality}
# Export Model 5 adjusted commonality
write.csv(m5_commonality_adjusted_final, "m5_neuro_age_commonality_adjusted.csv", row.names = FALSE)
cat("Exported Model 5 adjusted commonality to m5_neuro_age_commonality_adjusted.csv\n")

# Export Model 6 adjusted commonality
write.csv(m6_commonality_adjusted_final, "m6_hyperactivity_commonality_adjusted.csv", row.names = FALSE)
cat("Exported Model 6 adjusted commonality to m6_hyperactivity_commonality_adjusted.csv\n")

# Export Model 7 adjusted commonality
write.csv(m7_commonality_adjusted_final, "m7_inattention_commonality_adjusted.csv", row.names = FALSE)
cat("Exported Model 7 adjusted commonality to m7_inattention_commonality_adjusted.csv\n")
```


```{r Export for spaghetti plot symptoms}
# Export Data for Python Visualization
# Selects participants with >1 timepoint (already filtered in analysis_data)
# Includes raw variables and stable phenotype labels for plotting
# viz_export_data <- analysis_data %>%
#  select(
#    subject_id,
#    participant_id,
#    Age,             # Raw Age for X-axis
#    sex,             # Included in case you need to split/hue by sex
#    ADHDdiagnosis,   # The stable phenotype (Control/ADHD) for Hue
#    InattentionRaw,  # Raw Symptom Score
#    HyperactivityRaw,# Raw Symptom Score
#    true_g,          # Raw Cognition Score
#    stacked_pred     # Raw Model Prediction
#  )

# Write to CSV
# write.csv(viz_export_data, "longitudinal_phenotype_symptom_data.csv", row.names = FALSE)
#
#cat("Exported visualization dataset to 'longitudinal_phenotype_symptom_data.csv' with", nrow(viz_export_data), "rows.\n")
```

# Sensitivity Analysis: Load Covariate Data
```{r sensitivity_load_covariates}
# Loads and parses three external covariate sources:
#   1. Volumes remaining post-censoring (motion QC)
#   2. Euler numbers (structural surface QC)
#   3. Medication status (stimulant use)
#
# subject_id in volumes/euler files: sub-NDARINVXXXXXXX_Age120Months_StudyYear01
# Matched to analysis_data on participant_id (NDAR ID) and Age (months)
# Medication matched on NDAR ID with +/-6 month age tolerance

# -- Helper: parse XCP-D/FreeSurfer subject_id format --
parse_xcpd_id <- function(df, col = "subject_id") {
  df %>%
    mutate(
      stripped      = str_remove(.data[[col]], "^sub-"),
      participant_id = str_extract(stripped, "^.+?(?=Age)"),
      Age            = as.numeric(str_extract(stripped, "(?<=Age)\\d+(?=Months)"))
    ) %>%
    select(-stripped)
}

# -- Volumes remaining post-censoring --
volumes_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/Linear_mixed_effects_OREGON/volumes_censored_by_subject_summary.csv",
  show_col_types = FALSE
)

volumes_clean <- volumes_raw %>%
  parse_xcpd_id() %>%
  select(participant_id, Age, perc_remaining_post_censoring,
         volumes_remaining_post_censoring)

# -- Euler numbers (structural QC): mean across hemispheres --
euler_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/oregon_adhd_1000_euler_metrics_v3.csv",
  show_col_types = FALSE
)

euler_clean <- euler_raw %>%
  parse_xcpd_id() %>%
  mutate(mean_euler = (lh_euler + rh_euler) / 2) %>%
  select(participant_id, Age, mean_euler)

# -- Medication data --
# Values: 1 = no, 3 = yes; recoded to 0/1
# med_age stored separately to avoid column name conflict on join
med_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/meddem01.csv",
  show_col_types = FALSE
)

med_clean <- med_raw %>%
  mutate(
    participant_id      = str_remove_all(subjectkey, "_"),
    med_age             = as.numeric(interview_age),
    current_stimmed_bin = case_when(current_stimmed == 3 ~ 1L,
                                    current_stimmed == 1 ~ 0L,
                                    TRUE                 ~ NA_integer_),
    stimmed_bin         = case_when(stimmed == 3 ~ 1L,
                                    stimmed == 1 ~ 0L,
                                    TRUE         ~ NA_integer_)
  ) %>%
  select(participant_id, med_age, current_stimmed_bin, stimmed_bin) %>%
  filter(!is.na(current_stimmed_bin) | !is.na(stimmed_bin))

# -- Parental SES (one row per subject after deduplication) --
# Education values mapped to ordinal scale; highest of mother/father retained.
# ps_prime_income retained as-is after -999 removal.
edu_map <- c(
  "Less than High School" = 0,
  "Some High School"      = 1,
  "High School Degree"    = 2,
  "GED"                   = 2,
  "Some College"          = 3,
  "Associates Degree"     = 3,
  "Bachelors Degree"      = 3,
  "Graduate Degree"       = 3
)

socdem_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/socdem01.csv",
  show_col_types = FALSE
)

socdem_clean <- socdem_raw %>%
  mutate(participant_id = str_remove_all(subjectkey, "_")) %>%
  mutate(
    bio_mother_education = ifelse(bio_mother_education %in% c(-999, 77), NA, bio_mother_education),
    bio_father_education = ifelse(bio_father_education %in% c(-999, 77), NA, bio_father_education),
    ps_prime_income      = as.numeric(ifelse(ps_prime_income == -999, NA, ps_prime_income))
  ) %>%
  mutate(
    mother_edu  = recode(bio_mother_education, !!!edu_map, .default = NA_real_),
    father_edu  = recode(bio_father_education, !!!edu_map, .default = NA_real_),
    highest_edu = pmax(mother_edu, father_edu, na.rm = TRUE)
  ) %>%
  group_by(participant_id) %>%
  summarise(
    highest_edu     = first(na.omit(highest_edu)),
    ps_prime_income = first(na.omit(ps_prime_income)),
    .groups = "drop"
  )

cat("Covariate data loaded.\n")
cat(sprintf("  Volumes rows:    %d\n", nrow(volumes_clean)))
cat(sprintf("  Euler rows:      %d\n", nrow(euler_clean)))
cat(sprintf("  Medication rows: %d\n", nrow(med_clean)))
cat(sprintf("  Socdem rows:     %d\n", nrow(socdem_clean)))
```

# Sensitivity Analysis: Merge Covariates
```{r sensitivity_merge_covariates}
# Merges covariate data into analysis_data_full.
# Volumes and Euler matched on participant_id + Age (exact).
# Medication matched on participant_id + nearest age within +/-6 months.
# analysis_data is then updated from the merged analysis_data_full.

analysis_data_full <- analysis_data_full %>%
  left_join(volumes_clean, by = c("participant_id", "Age")) %>%
  left_join(euler_clean,   by = c("participant_id", "Age"))

# Fuzzy medication merge: closest record within +/-6 months
analysis_data_full <- analysis_data_full %>%
  left_join(med_clean, by = "participant_id", relationship = "many-to-many") %>%
  mutate(age_diff = abs(Age - med_age)) %>%
  filter(age_diff <= 6 | is.na(age_diff)) %>%
  group_by(subject_id) %>%
  slice_min(age_diff, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-med_age, -age_diff)

# Refresh longitudinal subset with merged covariates
analysis_data <- analysis_data_full %>%
  filter(count_timepoints >= 2)

# Socdem joins on participant_id directly (already deduplicated to one row per subject)
analysis_data_full <- analysis_data_full %>%
  left_join(socdem_clean, by = "participant_id")

analysis_data <- analysis_data_full %>%
  filter(count_timepoints >= 2)

cat("Covariate merge complete. Coverage in full dataset:\n")
cat(sprintf("  Volumes remaining:   %d / %d\n",
            sum(!is.na(analysis_data_full$perc_remaining_post_censoring)),
            nrow(analysis_data_full)))
cat(sprintf("  Euler number:        %d / %d\n",
            sum(!is.na(analysis_data_full$mean_euler)),
            nrow(analysis_data_full)))
cat(sprintf("  Current stimmed:     %d / %d\n",
            sum(!is.na(analysis_data_full$current_stimmed_bin)),
            nrow(analysis_data_full)))
cat(sprintf("  Highest edu:         %d / %d\n",
            sum(!is.na(analysis_data_full$highest_edu)),
            nrow(analysis_data_full)))
```


# Sensitivity Analysis: Sex Comparison
```{r sensitivity_sex_comparison power analysis}
# Brain-cognition relationship assessed separately for males and females,
# with a TOST equivalence test on the Fisher Z-transformed correlations.
# First, power analysis:

ALPHA <- 0.05
POWER <- 0.80

sens_sex_data <- analysis_data_full %>%
  filter(!is.na(sex), !is.na(true_g_z), !is.na(stacked_pred_z))

n_male_justify   <- sum(sens_sex_data$sex == "M")
n_female_justify <- sum(sens_sex_data$sex == "F")

se_diff_power <- sqrt(1 / (n_male_justify - 3) + 1 / (n_female_justify - 3))

z_alpha <- qnorm(1 - ALPHA / 2)
z_power <- qnorm(POWER)

q_detectable_80 <- se_diff_power * (z_alpha + z_power)

cat("=== Equivalence Bound Justification (resource-based SESOI) ===\n")
cat(sprintf("N: male = %d, female = %d\n", n_male_justify, n_female_justify))
cat(sprintf("SE of Fisher z difference: %.4f\n", se_diff_power))
cat(sprintf("Minimum Fisher z difference detectable with %.0f%% power at alpha = %.2f: %.4f\n",
            POWER * 100, ALPHA, q_detectable_80))

```

```{r set equivalence and do test}

# Equivalence bound: +/-0.2 in Fisher Z units (approximately equivalent to
# a difference of ~0.19 in r for correlations around 0.45).
# 
# Resource-based justification for the equivalence bound, following the
# worked example illustrated in Lakens et al. (2018), where the bound is set to
# the effect size the design had adequate power to detect, rather than a
# benchmark value. Computed here for the male vs female comparison prior
# to the equivalence test below.
#
# p_equiv < 0.05 indicates the male and female brain-cognition correlations
# are statistically equivalent within the specified bounds.

EQUIV_BOUND <- 0.1786

sens_sex_data <- analysis_data_full %>%
  filter(!is.na(sex), !is.na(true_g_z), !is.na(stacked_pred_z))

data_male   <- sens_sex_data %>% filter(sex == "M")
data_female <- sens_sex_data %>% filter(sex == "F")

m00_male   <- lm(true_g_z ~ stacked_pred_z, data = data_male)
m00_female <- lm(true_g_z ~ stacked_pred_z, data = data_female)

r_male   <- cor(data_male$stacked_pred_z,   data_male$true_g_z,   use = "complete.obs")
r_female <- cor(data_female$stacked_pred_z, data_female$true_g_z, use = "complete.obs")

n_male   <- sum(!is.na(data_male$stacked_pred_z)   & !is.na(data_male$true_g_z))
n_female <- sum(!is.na(data_female$stacked_pred_z) & !is.na(data_female$true_g_z))

cat("=== Sex Sensitivity: Brain-Cognition Relationship by Sex ===\n")
cat(sprintf("Male:    r = %.3f,  R2 = %.4f,  N = %d\n",
            r_male,   summary(m00_male)$r.squared,   n_male))
cat(sprintf("Female:  r = %.3f,  R2 = %.4f,  N = %d\n",
            r_female, summary(m00_female)$r.squared, n_female))

# -- TOST equivalence test on Fisher Z-transformed correlations --
z_male   <- atanh(r_male)
z_female <- atanh(r_female)
z_diff   <- z_male - z_female
se_diff  <- sqrt(1 / (n_male - 3) + 1 / (n_female - 3))
df_tost  <- n_male + n_female - 4

t_lower  <- (z_diff - (-EQUIV_BOUND)) / se_diff
t_upper  <- (z_diff -   EQUIV_BOUND)  / se_diff
p_lower  <- pt(t_lower, df = df_tost, lower.tail = FALSE)
p_upper  <- pt(t_upper, df = df_tost, lower.tail = TRUE)
p_equiv  <- max(p_lower, p_upper)

# Conventional difference test for reference
z_se_null <- sqrt(1 / (n_male - 3) + 1 / (n_female - 3))
z_stat    <- z_diff / z_se_null
p_diff    <- 2 * pnorm(abs(z_stat), lower.tail = FALSE)

cat(sprintf("\nFisher Z: male = %.4f, female = %.4f, difference = %.4f\n",
            z_male, z_female, z_diff))
cat(sprintf("SE of difference: %.4f\n", se_diff))
cat(sprintf("\nEquivalence bounds: +/- %.2f Fisher Z units\n", EQUIV_BOUND))
cat(sprintf("t_lower = %.4f,  p_lower = %.4f\n", t_lower, p_lower))
cat(sprintf("t_upper = %.4f,  p_upper = %.4f\n", t_upper, p_upper))
cat(sprintf("p_equiv = %.4f  (max of p_lower and p_upper)\n", p_equiv))
if (p_equiv < 0.05) {
  cat("Result: correlations are statistically equivalent within bounds.\n")
} else {
  cat("Result: equivalence within bounds not established.\n")
}

cat(sprintf("\nConventional difference test: z = %.4f, p = %.4f\n", z_stat, p_diff))

cat("\nMale model summary:\n")
print(summary(m00_male))

cat("\nFemale model summary:\n")
print(summary(m00_female))


age_by_sex <- sens_sex_data %>%
  group_by(sex) %>%
  summarise(
    n = n(),
    mean_age = mean(Age, na.rm = TRUE),
    sd_age = sd(Age, na.rm = TRUE),
    min_age = min(Age, na.rm = TRUE),
    max_age = max(Age, na.rm = TRUE)
  )

print(age_by_sex)

# Welch's t-test as a simple descriptive check for whether age differs
# by sex in this subset.
age_test <- t.test(Age ~ sex, data = sens_sex_data)
print(age_test)


# Triple check our N

analysis_data_full %>% filter(!is.na(sex)) %>% summarise(n(), n_distinct(participant_id))
```

