---
title: "LME_Oregon_Cog_and_symptoms_AIO"
author: "Jack Scott"
date: "2026-01-23"
output:
  html_document:
    df_print: paged
---

# Setup
```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir ='/media/hcs-sci-psy-narun/Jack/First_level_model_aligned/Stacked_model_ridge_v4_EXTREME_conserve_a2000_keeper')
``` 

# Packages
```{r packages}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(r2mlm)
library(glmm.hp)
library(broom.mixed)
```

# Data Preparation
```{r data_prep}
# Load Neuroimaging/Cognition Data
cognition_data <- readr::read_csv("/media/hcs-sci-psy-narun/Jack/First_level_model_aligned/Stacked_model_ridge_v4_EXTREME_conserve_a2000_keeper/stacked_ridge_predictions.csv")

# Load Phenotype Data (ADHD Diagnosis)
phenotype_data <- read_csv("/media/hcs-sci-psy-narun/Jack/HMM_ELM_data_prep/phase4_statistical_analysis/analysis_ready_dataset.csv")

# Load Inattention/Hyperactivity Matched Datasets
inattention_scores <- readr::read_csv("/media/hcs-sci-psy-narun/Jack/LME_cog_inattention/inattention_scores_matched.csv")
hyperactivity_scores <- readr::read_csv("/media/hcs-sci-psy-narun/Jack/LME_cog_hyperactivity/hyperactivity_scores_matched.csv")

# Pre-processing & Merging
analysis_data <- cognition_data %>%
  # Extract metadata
  mutate(participant_id = str_extract(subject_id, "^[^_]+"),
         Age = as.numeric(str_extract(subject_id, "(?<=_Age)\\d+"))) %>%
  # Merge Phenotypes
  left_join(
    phenotype_data %>% 
      mutate(subject_id_match = paste0(subject_id, "_Age", age_months)) %>%
      select(subject_id_match, final_phenotype, sex),
    by = c("subject_id" = "subject_id_match")
  ) %>%
  # Clean Diagnosis Label
  mutate(ADHDdiagnosis = factor(case_when(
    final_phenotype == 3 ~ "ADHD",
    final_phenotype == 1 ~ "Control",
    TRUE ~ NA_character_
  ), levels = c("Control", "ADHD"))) %>%
  # Merge Symptoms
  mutate(
    InattentionRaw = inattention_scores$inattention[match(subject_id, inattention_scores$subject_id)],
    HyperactivityRaw = hyperactivity_scores$hyperactivity[match(subject_id, hyperactivity_scores$subject_id)]
  ) %>%
  # Standardizing
  mutate(
    participant_id = as_factor(participant_id),
    Age0 = Age - min(Age, na.rm = TRUE),
    true_g_z = as.double(scale(true_g)),
    stacked_pred_z = as.double(scale(stacked_pred)),
    inattention_z = as.double(scale(InattentionRaw)),
    hyperactivity_z = as.double(scale(HyperactivityRaw))
  ) %>%
  # Within/Between Decomposition
  group_by(participant_id) %>%
    mutate(
      true_g_z_mean = mean(true_g_z, na.rm = TRUE),
      stacked_pred_z_mean = mean(stacked_pred_z, na.rm = TRUE),
      InattentionMean = mean(inattention_z, na.rm = TRUE),
      HyperactivityMean = mean(hyperactivity_z, na.rm = TRUE),
      age_mean = mean(Age0, na.rm = TRUE),
      count_timepoints = n()
    ) %>%
  ungroup() %>%
  mutate(
    true_g_z_dev = true_g_z - true_g_z_mean,
    stacked_pred_z_dev = stacked_pred_z - stacked_pred_z_mean,
    InattentionDeviation = inattention_z - InattentionMean,
    HyperactivityDeviation = hyperactivity_z - HyperactivityMean,
    age_deviation = Age0 - age_mean
  ) %>%
  # Filter for longitudinal analysis compatibility
  filter(count_timepoints >= 2)

cat("Final analysis dataset size:", nrow(analysis_data), "observations from", n_distinct(analysis_data$participant_id), "participants.\n")
```

```{r}
# Calculate Pearson correlation between deviation scores
pearson_correlation_deviations <- cor(analysis_data$true_g_z_dev, 
                          analysis_data$stacked_pred_z_dev, 
                          method = "pearson",
                          use = "complete.obs")

# Get unique participant-level data (since means are repeated for each observation)
participant_means <- analysis_data[!duplicated(analysis_data$participant_id), ]

# Create a scatterplot to visualize the relationship
library(ggplot2)
ggplot(analysis_data, aes(x = true_g_z_dev, y = stacked_pred_z_dev)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "True g (deviation scores)", 
       y = "Predicted g (deviation scores)",
       title = "Within-Person Relationship: Observed vs Predicted g",
       subtitle = paste("Pearson r =", round(pearson_correlation_deviations, 3))) +
  theme_minimal()

# Calculate Pearson correlation between participant means
pearson_means <- cor(participant_means$true_g_z_mean, 
                    participant_means$stacked_pred_z_mean, 
                    method = "pearson",
                    use = "complete.obs")

# Create a scatterplot for participant means
ggplot(participant_means, aes(x = true_g_z_mean, y = stacked_pred_z_mean)) +
  geom_point(alpha = 0.7, size = 2, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red", fill = "red", alpha = 0.2) +  
  # Add correlation statistics in a properly sized box (lower right)
  annotate("label", 
           x = max(participant_means$true_g_z_mean, na.rm = TRUE) * 0.95, 
           y = min(participant_means$stacked_pred_z_mean, na.rm = TRUE) * 0.95,
           label = paste("Pearson r =", round(pearson_means, 3), 
                         "\nn =", nrow(participant_means)),
           size = 3,
           fill = "white", 
           color = "black",
           alpha = 0.9,
           label.size = 0.15) +  
  labs(x = "Observed g (participant means)", 
       y = "Predicted g (participant means)",
       title = "Between-Person Relationship: Observed vs Predicted g") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        # Note: Changed size to linewidth to avoid ggplot2 deprecation warnings
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5))
``` 


# Model 1: ADHD Additive
```{r m1_adhd_additive}
m1_adhd_add <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + ADHDdiagnosis + (1 | participant_id), 
                    data = analysis_data, REML = FALSE)
summary(m1_adhd_add)
```

# Model 2: ADHD Interactive
```{r m2_adhd_interactive}
m2_adhd_int <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + ADHDdiagnosis + 
                    stacked_pred_z_dev:ADHDdiagnosis + stacked_pred_z_mean:ADHDdiagnosis + (1 | participant_id), 
                    data = analysis_data, REML = FALSE)
summary(m2_adhd_int)
```

# Model 3: Neuroimaging-only
```{r m3_neuroimaging}
m3_neuro <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m3_neuro)
```

# Model 4: Age-only
```{r m4_age}
m4_age <- lmer(true_g_z ~ age_deviation + age_mean + (1 | participant_id), 
               data = analysis_data, REML = FALSE)
summary(m4_age)
```

# Model 5: Neuroimaging-and-age
```{r m5_neuro_age}
m5_neuro_age <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + age_deviation + age_mean + (1 | participant_id), 
                     data = analysis_data, REML = FALSE)
summary(m5_neuro_age)

print("\n ===DEBUG===")
# Diagnostic: Print full r2mlm structure
# r2_full <- r2mlm::r2mlm(m5_neuro_age)
# print("Full r2mlm output:")
# print(r2_full)
# print("\nNames in r2mlm object:")
# print(names(r2_full))
# print("\nStructure:")
# str(r2_full)

```

# Model 6: Hyperactivity
```{r m6_hyperactivity}
m6_hyper <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + HyperactivityDeviation + HyperactivityMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m6_hyper)
```
# hyperactivity only
```{r m6p5_hyperactivity_only}
m6p5_hyper <- lmer(true_g_z ~ HyperactivityDeviation + HyperactivityMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m6p5_hyper)
```


# Model 7: Inattention
```{r m7_inattention}
m7_inatt <- lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + InattentionDeviation + InattentionMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m7_inatt)
```

```{r m7p5_inattention}
m7p5_inatt <- lmer(true_g_z ~ InattentionDeviation + InattentionMean + (1 | participant_id), 
                 data = analysis_data, REML = FALSE)
summary(m7p5_inatt)
```

# ADHD Model Comparison
```{r adhd_comparison}
# Formal comparison
comp_adhd <- anova(m1_adhd_add, m2_adhd_int)
print(comp_adhd)

# BIC Difference logic
delta_bic <- BIC(m2_adhd_int) - BIC(m1_adhd_add)
cat("\nDelta BIC (Interactive - Additive):", round(delta_bic, 2), "\n")
if(delta_bic > 0) {
  cat("Evidence favors the simpler ADDITIVE model.\n")
} else {
  cat("Evidence favors the INTERACTIVE model.\n")
}
```

# Core Models Comparison
```{r core_comparison}
comp_biomarkers_age <- anova(m3_neuro, m4_age, m5_neuro_age)
print(comp_biomarkers_age)
```

# SJPlot Tables: Core Models
```{r sjplot_core}
tab_model(m3_neuro, m4_age, m5_neuro_age, 
          dv.labels = c("Neuroimaging Only", "Age Only", "Neuroimaging-and-Age"),
          pred.labels = c(
            "(Intercept)" = "Intercept",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "age_deviation" = "age<sub>deviation</sub>",
            "age_mean" = "age<sub>mean</sub>"
          ),
          show.aic = TRUE,
          show.dev = TRUE,  # Shows deviance instead of BIC
          title = "Model Comparison: Neuroimaging and Age",
          file = "core_models")
```

# SJPlot Tables: ADHD Models
```{r sjplot_adhd}
tab_model(m1_adhd_add, m2_adhd_int, 
          dv.labels = c("ADHD Additive", "ADHD Interactive"),
          pred.labels = c(
            "(Intercept)" = "(Intercept)",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD" = "ADHD Diagnosis",
            "stacked_pred_z_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD"),
          show.aic = TRUE,
          show.dev = TRUE,
          title = "ADHD Diagnosis Comparison",
          file = "adhd_diagnosis")
```

# SJPlot Tables: Symptom models
```{r sjplot_hyp}
tab_model(m6p5_hyper, m6_hyper, 
          dv.labels = c("Hyperactivity only", "Neuroimaging-and-Hyperactivity"),
          pred.labels = c(
            "(Intercept)" = "Intercept",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD" = "ADHD Diagnosis",
            "stacked_pred_z_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD",
            "HyperactivityMean" = "Hyperactivity<sub>mean</sub>",
            "HyperactivityDeviation" = "Hyperactivity<sub>deviation</sub>"
          ),
          show.aic = TRUE,
          show.dev = TRUE,
          title = "Symptoms Comparison",
          file = "symptom_models")
```

```{r sjplot_inatt}
tab_model(m7p5_inatt, m7_inatt, 
          dv.labels = c("Inattention Only", "Neuroimaging-and-Inattention"),
          pred.labels = c(
            "(Intercept)" = "Intercept",
            "stacked_pred_z_dev" = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_mean" = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD" = "ADHD Diagnosis",
            "stacked_pred_z_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD",
            "InattentionMean" = "Inattention<sub>mean</sub>",
            "InattentionDeviation" = "Inattention<sub>deviation</sub>" 
          ),
          show.aic = TRUE,
          show.dev = TRUE,
          title = "Symptoms Comparison",
          file = "symptom_models")
```

# R2 Decomposition: Model 5
```{r m5_r2_calculation}
# Calculate R-squared for Model 5
r2_m5 <- r2mlm::r2mlm(m5_neuro_age)
print("Model 5 R-squared:")
print(r2_m5)
```

# Commonality Analysis: Model 5
```{r m5_commonality}
# Refit using lme4 for glmm.hp compatibility
m5_lme4 <- lme4::lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + age_deviation + age_mean + (1 | participant_id), 
                      data = analysis_data, REML = FALSE)

# Run commonality analysis
m5_commonality_results <- glmm.hp::glmm.hp(m5_lme4, type = "adjR2", commonality = TRUE)
print("Model 5 Commonality Results:")
print(m5_commonality_results)
```

# Adjust Commonality: Model 5
```{r m5_adjust_commonality}
# Extract commonality analysis results
m5_commonality_raw <- m5_commonality_results$commonality.analysis

# Remove total rows
is_total_row <- grepl("total|Total|TOTAL", rownames(m5_commonality_raw), ignore.case = TRUE)
m5_components_only <- m5_commonality_raw[!is_total_row, , drop = FALSE]

# Set negative values to zero
m5_components_positive <- ifelse(m5_components_only[,1] < 0, 0, m5_components_only[,1])
names(m5_components_positive) <- rownames(m5_components_only)

# Re-normalize to 100%
m5_sum_positive <- sum(m5_components_positive)
m5_components_adjusted_pct <- (m5_components_positive / m5_sum_positive) * 100

# Create adjusted results table
m5_commonality_adjusted <- data.frame(
  Component = rownames(m5_components_only),
  Original_Fraction = m5_components_only[,1],
  Original_Percent = m5_components_only[,2],
  Adjusted_Fraction = m5_components_positive,
  Adjusted_Percent = m5_components_adjusted_pct,
  stringsAsFactors = FALSE
)

# Add adjusted total row
m5_total_row <- data.frame(
  Component = "Total",
  Original_Fraction = sum(m5_components_only[,1]),
  Original_Percent = 100.00,
  Adjusted_Fraction = m5_sum_positive,
  Adjusted_Percent = 100.00,
  stringsAsFactors = FALSE
)

# Combine
m5_commonality_adjusted_final <- rbind(m5_commonality_adjusted, m5_total_row)
print("Model 5 Adjusted Commonality:")
print(m5_commonality_adjusted_final)
```

# Commonality Analysis: Model 6 (Hyperactivity)
```{r m6_commonality}
# Refit using lme4
m6_lme4 <- lme4::lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + HyperactivityDeviation + HyperactivityMean + (1 | participant_id), 
                      data = analysis_data, REML = FALSE)

# Run commonality analysis
m6_commonality_results <- glmm.hp::glmm.hp(m6_lme4, type = "adjR2", commonality = TRUE)
print("Model 6 Commonality Results:")
print(m6_commonality_results)
```

# Adjust Commonality: Model 6
```{r m6_adjust_commonality}
# Extract commonality analysis results
m6_commonality_raw <- m6_commonality_results$commonality.analysis

# Remove total rows
is_total_row <- grepl("total|Total|TOTAL", rownames(m6_commonality_raw), ignore.case = TRUE)
m6_components_only <- m6_commonality_raw[!is_total_row, , drop = FALSE]

# Set negative values to zero
m6_components_positive <- ifelse(m6_components_only[,1] < 0, 0, m6_components_only[,1])
names(m6_components_positive) <- rownames(m6_components_only)

# Re-normalize to 100%
m6_sum_positive <- sum(m6_components_positive)
m6_components_adjusted_pct <- (m6_components_positive / m6_sum_positive) * 100

# Create adjusted results table
m6_commonality_adjusted <- data.frame(
  Component = rownames(m6_components_only),
  Original_Fraction = m6_components_only[,1],
  Original_Percent = m6_components_only[,2],
  Adjusted_Fraction = m6_components_positive,
  Adjusted_Percent = m6_components_adjusted_pct,
  stringsAsFactors = FALSE
)

# Add adjusted total row
m6_total_row <- data.frame(
  Component = "Total",
  Original_Fraction = sum(m6_components_only[,1]),
  Original_Percent = 100.00,
  Adjusted_Fraction = m6_sum_positive,
  Adjusted_Percent = 100.00,
  stringsAsFactors = FALSE
)

# Combine
m6_commonality_adjusted_final <- rbind(m6_commonality_adjusted, m6_total_row)
print("Model 6 Adjusted Commonality:")
print(m6_commonality_adjusted_final)
```

# Commonality Analysis: Model 7 (Inattention)
```{r m7_commonality}
# Refit using lme4
m7_lme4 <- lme4::lmer(true_g_z ~ stacked_pred_z_dev + stacked_pred_z_mean + InattentionDeviation + InattentionMean + (1 | participant_id), 
                      data = analysis_data, REML = FALSE)

# Run commonality analysis
m7_commonality_results <- glmm.hp::glmm.hp(m7_lme4, type = "adjR2", commonality = TRUE)
print("Model 7 Commonality Results:")
print(m7_commonality_results)
```

# Adjust Commonality: Model 7
```{r m7_adjust_commonality}
# Extract commonality analysis results
m7_commonality_raw <- m7_commonality_results$commonality.analysis

# Remove total rows
is_total_row <- grepl("total|Total|TOTAL", rownames(m7_commonality_raw), ignore.case = TRUE)
m7_components_only <- m7_commonality_raw[!is_total_row, , drop = FALSE]

# Set negative values to zero
m7_components_positive <- ifelse(m7_components_only[,1] < 0, 0, m7_components_only[,1])
names(m7_components_positive) <- rownames(m7_components_only)

# Re-normalize to 100%
m7_sum_positive <- sum(m7_components_positive)
m7_components_adjusted_pct <- (m7_components_positive / m7_sum_positive) * 100

# Create adjusted results table
m7_commonality_adjusted <- data.frame(
  Component = rownames(m7_components_only),
  Original_Fraction = m7_components_only[,1],
  Original_Percent = m7_components_only[,2],
  Adjusted_Fraction = m7_components_positive,
  Adjusted_Percent = m7_components_adjusted_pct,
  stringsAsFactors = FALSE
)

# Add adjusted total row
m7_total_row <- data.frame(
  Component = "Total",
  Original_Fraction = sum(m7_components_only[,1]),
  Original_Percent = 100.00,
  Adjusted_Fraction = m7_sum_positive,
  Adjusted_Percent = 100.00,
  stringsAsFactors = FALSE
)

# Combine
m7_commonality_adjusted_final <- rbind(m7_commonality_adjusted, m7_total_row)
print("Model 7 Adjusted Commonality:")
print(m7_commonality_adjusted_final)
```

# Export R2 Decomposition: Model 5
```{r export_m5_r2_decomposition}
# Export R2 Decomposition: Model 5 - CORRECTED
# Preserve the matrix structure (total/within/between columns)

r2_decomposition <- r2mlm::r2mlm(m5_neuro_age)

# Extract the Decompositions matrix
decomp_matrix <- r2_decomposition$Decompositions

# Create dataframe with proper row and column labels
decomp_df <- data.frame(
  component = rownames(decomp_matrix),
  total = decomp_matrix[, "total"],
  within = decomp_matrix[, "within"],
  between = decomp_matrix[, "between"],
  stringsAsFactors = FALSE
)

# Export for Python
write.csv(decomp_df, "m5_r2mlm_decomposition.csv", row.names = FALSE)
cat("Exported Model 5 R2 decomposition to m5_r2mlm_decomposition.csv\n")
```

# Export Adjusted Commonality: All Models
```{r export_commonality}
# Export Model 5 adjusted commonality
write.csv(m5_commonality_adjusted_final, "m5_neuro_age_commonality_adjusted.csv", row.names = FALSE)
cat("Exported Model 5 adjusted commonality to m5_neuro_age_commonality_adjusted.csv\n")

# Export Model 6 adjusted commonality
write.csv(m6_commonality_adjusted_final, "m6_hyperactivity_commonality_adjusted.csv", row.names = FALSE)
cat("Exported Model 6 adjusted commonality to m6_hyperactivity_commonality_adjusted.csv\n")

# Export Model 7 adjusted commonality
write.csv(m7_commonality_adjusted_final, "m7_inattention_commonality_adjusted.csv", row.names = FALSE)
cat("Exported Model 7 adjusted commonality to m7_inattention_commonality_adjusted.csv\n")
```

```{r hyperactivity_commonality_viz_filled}
# Hyperactivity Variance Decomposition - Four-way Euler Structure
library(eulerr)

# -------------------------------------------------------------------------
# DEFINE DATA
# -------------------------------------------------------------------------
# Data extracted from m6_hyperactivity_commonality_adjusted.csv (Adjusted_Percent)
# "Within"  = stacked_pred_z_dev
# "Between" = stacked_pred_z_mean
# "HypDev"  = HyperactivityDeviation
# "HypMean" = HyperactivityMean

variance_data_4way <- c(
  # 1. Unique Variances
  "Within"  = 13.83,
  "Between" = 47.04,
  "HypDev"  = 0.00,   # Adjusted unique variance is 0 (fully shared)
  "HypMean" = 15.97,

  # 2. Pairwise Intersections
  "Between&Within"  = 0.21,
  "HypDev&Within"   = 4.45,
  "Between&HypDev"  = 6.09,
  "HypMean&Within"  = 0.11,
  "Between&HypMean" = 12.22,
  "HypDev&HypMean"  = 0.00,

  # 3. Three-way Intersections
  "Between&HypDev&Within"  = 0.00,
  "Between&HypMean&Within" = 0.00,
  "HypDev&HypMean&Within"  = 0.04,
  "Between&HypDev&HypMean" = 0.00,

  # 4. Four-way Intersection
  "Between&HypDev&HypMean&Within" = 0.04
)

print("Hyperactivity four-way variance data:")
print(variance_data_4way)

# -------------------------------------------------------------------------
# GENERATE EULER DIAGRAM
# -------------------------------------------------------------------------
# shape = "ellipse" allows the geometry to distort to fit the 0 variance subsets
euler_fit_4way <- euler(variance_data_4way, 
                        shape = "ellipse", 
                        control = list(extraopt = TRUE))

print(paste("Fit stress:", round(euler_fit_4way$stress, 4)))
print(paste("Diag Error:", round(euler_fit_4way$diagError, 4)))

# -------------------------------------------------------------------------
# PLOT
# -------------------------------------------------------------------------
p_euler_4way <- plot(euler_fit_4way,
     quantities = FALSE, #list(type = "percent", font = 2, round = 1),
     # Adjust colors as preferred
     fills = c("lightblue", "lightcoral", "lightgreen", "lightyellow"), 
     edges = list(col = "black", lwd = 1.0),
     labels = FALSE, #list(font = 2), 
     legend = TRUE,
     main = "Hyperactivity Variance Decomposition")

print(p_euler_4way)
``` 
```{r euler for inattention}
# Inattention Variance Decomposition - Four-way Euler Structure
library(eulerr)


# DEFINE DATA

variance_data_inatt <- c(
  # 1. Unique Variances
  "Within"    = 10.39,
  "Between"   = 36.12,
  "InattDev"  = 0.00,   # Unique variance is 0
  "InattMean" = 33.03,

  # 2. Pairwise Intersections
  "Between&Within"    = 0.08,
  "InattDev&Within"   = 3.90,
  "Between&InattDev"  = 4.86,
  "InattMean&Within"  = 1.01,
  "Between&InattMean" = 10.59,
  "InattDev&InattMean"= 0.00,

  # 3. Three-way Intersections (All effectively 0 in this dataset)
  "Between&InattDev&Within"    = 0.00,
  "Between&InattMean&Within"   = 0.00,
  "InattDev&InattMean&Within"  = 0.00,
  "Between&InattDev&InattMean" = 0.00,

  # 4. Four-way Intersection
  "Between&InattDev&InattMean&Within" = 0.00
)

print("Inattention four-way variance data:")
print(variance_data_inatt)

# -------------------------------------------------------------------------
# GENERATE EULER DIAGRAM
# -------------------------------------------------------------------------
# shape = "ellipse" is used to try and handle the disjoint/0-variance structure
euler_fit_inatt <- euler(variance_data_inatt, 
                         shape = "ellipse", 
                         control = list(extraopt = TRUE))

print(paste("Fit stress:", round(euler_fit_inatt$stress, 4)))
print(paste("Diag Error:", round(euler_fit_inatt$diagError, 4)))

# -------------------------------------------------------------------------
# PLOT
# -------------------------------------------------------------------------
p_euler_inatt <- plot(euler_fit_inatt,
     quantities = FALSE, #list(type = "percent", font = 2, round = 1),
     # Adjust colors here
     fills = c("lightblue", "lightcoral", "lightgreen", "lightyellow"), 
     edges = list(col = "black", lwd = 1.0),
     labels = FALSE, #list(font = 2), 
     legend = TRUE,
     main = "Inattention Variance Decomposition")

print(p_euler_inatt)

``` 
```{r Export for spaghetti plot symptoms}
# Export Data for Python Visualization
# Selects participants with >1 timepoint (already filtered in analysis_data)
# Includes raw variables and stable phenotype labels for plotting
viz_export_data <- analysis_data %>%
  select(
    subject_id,
    participant_id,
    Age,             # Raw Age for X-axis
    sex,             # Included in case you need to split/hue by sex
    ADHDdiagnosis,   # The stable phenotype (Control/ADHD) for Hue
    InattentionRaw,  # Raw Symptom Score
    HyperactivityRaw,# Raw Symptom Score
    true_g,          # Raw Cognition Score
    stacked_pred     # Raw Model Prediction
  )

# Write to CSV
write.csv(viz_export_data, "longitudinal_phenotype_symptom_data.csv", row.names = FALSE)

cat("Exported visualization dataset to 'longitudinal_phenotype_symptom_data.csv' with", nrow(viz_export_data), "rows.\n")
```
