---
title: "LME_Residualised_Sensitivity"
author: "Jack Scott"
date: "2026"
output:
  html_document:
    df_print: paged
---

# Setup
```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = '/media/hcs-sci-psy-narun/Jack/First_level_model_aligned/Stacked_model_ridge_v4_EXTREME_conserve_a2000_keeper')
```

# Packages
```{r packages}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(r2mlm)
library(glmm.hp)
library(broom.mixed)
```

# Data Loading
```{r data_load}
# -- Core neuroimaging/cognition data --
cognition_data <- readr::read_csv(
  "/media/hcs-sci-psy-narun/Jack/First_level_model_aligned/Stacked_model_ridge_v4_EXTREME_conserve_a2000_keeper/stacked_ridge_predictions.csv")

phenotype_data <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/HMM_ELM_data_prep/phase4_statistical_analysis/analysis_ready_dataset.csv")

inattention_scores   <- readr::read_csv(
  "/media/hcs-sci-psy-narun/Jack/LME_cog_inattention/inattention_scores_matched.csv")
hyperactivity_scores <- readr::read_csv(
  "/media/hcs-sci-psy-narun/Jack/LME_cog_hyperactivity/hyperactivity_scores_matched.csv")

# -- Residualisation covariates --

# Volumes remaining post-censoring
parse_xcpd_id <- function(df, col = "subject_id") {
  df %>%
    mutate(
      stripped      = str_remove(.data[[col]], "^sub-"),
      participant_id = str_extract(stripped, "^.+?(?=Age)"),
      Age            = as.numeric(str_extract(stripped, "(?<=Age)\\d+(?=Months)"))
    ) %>%
    select(-stripped)
}

volumes_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/Linear_mixed_effects_OREGON/volumes_censored_by_subject_summary.csv",
  show_col_types = FALSE)

volumes_clean <- volumes_raw %>%
  parse_xcpd_id() %>%
  select(participant_id, Age, perc_remaining_post_censoring)

# Medication
med_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/meddem01.csv",
  show_col_types = FALSE)

med_clean <- med_raw %>%
  mutate(
    participant_id      = str_remove_all(subjectkey, "_"),
    med_age             = as.numeric(interview_age),
    current_stimmed_bin = case_when(current_stimmed == 3 ~ 1L,
                                    current_stimmed == 1 ~ 0L,
                                    TRUE                 ~ NA_integer_)
  ) %>%
  select(participant_id, med_age, current_stimmed_bin) %>%
  filter(!is.na(current_stimmed_bin))

# Euler number (structural surface quality: mean across hemispheres)
euler_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/oregon_adhd_1000_euler_metrics_v3.csv",
  show_col_types = FALSE)

euler_clean <- euler_raw %>%
  parse_xcpd_id() %>%
  mutate(mean_euler = (lh_euler + rh_euler) / 2) %>%
  select(participant_id, Age, mean_euler)

# Parental education (highest of mother/father, ordinal)
edu_map <- c(
  "Less than High School" = 0,
  "Some High School"      = 1,
  "High School Degree"    = 2,
  "GED"                   = 2,
  "Some College"          = 3,
  "Associates Degree"     = 3,
  "Bachelors Degree"      = 3,
  "Graduate Degree"       = 3
)

socdem_raw <- read_csv(
  "/media/hcs-sci-psy-narun/Jack/socdem01.csv",
  show_col_types = FALSE)

socdem_clean <- socdem_raw %>%
  mutate(participant_id = str_remove_all(subjectkey, "_")) %>%
  mutate(
    bio_mother_education = ifelse(bio_mother_education %in% c(-999, 77), NA, bio_mother_education),
    bio_father_education = ifelse(bio_father_education %in% c(-999, 77), NA, bio_father_education)
  ) %>%
  mutate(
    mother_edu  = recode(bio_mother_education, !!!edu_map, .default = NA_real_),
    father_edu  = recode(bio_father_education, !!!edu_map, .default = NA_real_),
    highest_edu = pmax(mother_edu, father_edu, na.rm = TRUE)
  ) %>%
  group_by(participant_id) %>%
  summarise(highest_edu = first(na.omit(highest_edu)), .groups = "drop")

cat("Data loaded.\n")
```

# Data Preparation
```{r data_prep}
analysis_data_full <- cognition_data %>%
  mutate(
    participant_id = str_extract(subject_id, "^[^_]+"),
    Age            = as.numeric(str_extract(subject_id, "(?<=_Age)\\d+"))
  ) %>%
  left_join(
    phenotype_data %>%
      mutate(subject_id_match = paste0(subject_id, "_Age", age_months)) %>%
      select(subject_id_match, final_phenotype, sex),
    by = c("subject_id" = "subject_id_match")
  ) %>%
  mutate(ADHDdiagnosis = factor(case_when(
    final_phenotype == 3 ~ "ADHD",
    final_phenotype == 1 ~ "Control",
    TRUE ~ NA_character_
  ), levels = c("Control", "ADHD"))) %>%
  mutate(
    InattentionRaw    = inattention_scores$inattention[match(subject_id, inattention_scores$subject_id)],
    HyperactivityRaw  = hyperactivity_scores$hyperactivity[match(subject_id, hyperactivity_scores$subject_id)]
  ) %>%
  mutate(
    participant_id   = as_factor(participant_id),
    Age0             = Age - min(Age, na.rm = TRUE),
    true_g_z         = as.double(scale(true_g)),
    stacked_pred_z   = as.double(scale(stacked_pred)),
    inattention_z    = as.double(scale(InattentionRaw)),
    hyperactivity_z  = as.double(scale(HyperactivityRaw))
  ) %>%
  # Merge covariates (exact match on participant_id + Age)
  left_join(volumes_clean, by = c("participant_id", "Age")) %>%
  # Medication fuzzy match: nearest record within +/-6 months
  left_join(med_clean, by = "participant_id", relationship = "many-to-many") %>%
  mutate(age_diff = abs(Age - med_age)) %>%
  filter(age_diff <= 6 | is.na(age_diff)) %>%
  group_by(subject_id) %>%
  slice_min(age_diff, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-med_age, -age_diff) %>%
  # Parental education (time-invariant, join on participant_id)
  left_join(socdem_clean, by = "participant_id") %>%
  # Euler number (timepoint-level, join on participant_id + Age)
  left_join(euler_clean %>% mutate(participant_id = as.character(participant_id)),
            by = c("participant_id", "Age")) %>%
  # Within/between decomposition and timepoint counting on full dataset
  group_by(participant_id) %>%
  mutate(
    true_g_z_mean        = mean(true_g_z, na.rm = TRUE),
    stacked_pred_z_mean  = mean(stacked_pred_z, na.rm = TRUE),
    InattentionMean      = mean(inattention_z, na.rm = TRUE),
    HyperactivityMean    = mean(hyperactivity_z, na.rm = TRUE),
    age_mean             = mean(Age0, na.rm = TRUE),
    count_timepoints     = n()
  ) %>%
  ungroup() %>%
  mutate(
    true_g_z_dev         = true_g_z - true_g_z_mean,
    stacked_pred_z_dev   = stacked_pred_z - stacked_pred_z_mean,
    InattentionDeviation = inattention_z - InattentionMean,
    HyperactivityDeviation = hyperactivity_z - HyperactivityMean,
    age_deviation        = Age0 - age_mean
  )

cat("Full dataset:", nrow(analysis_data_full), "observations from",
    n_distinct(analysis_data_full$participant_id), "participants.\n")
```

# Residualisation
```{r residualisation}
# Residualises stacked_pred_z from three confounds via OLS.
# true_g_z is left unchanged, outcome static
# Timepoints with missing confound data are excluded.
# count_timepoints is recomputed after confound-complete filtering.

# Filter to complete confound cases
analysis_data_resid <- analysis_data_full %>%
  filter(
    !is.na(current_stimmed_bin),
    !is.na(perc_remaining_post_censoring),
    !is.na(highest_edu),
    !is.na(mean_euler)
  )

cat("Timepoints after confound-complete filter:",
    nrow(analysis_data_resid), "\n")

# OLS residualisation of stacked predictions
resid_model <- lm(
  stacked_pred_z ~ current_stimmed_bin + perc_remaining_post_censoring + highest_edu + mean_euler,
  data = analysis_data_resid
)

cat("\nResidualisation model summary:\n")
print(summary(resid_model))

# Extract residuals and z-standardise
analysis_data_resid <- analysis_data_resid %>%
  mutate(stacked_pred_z_resid = as.double(scale(residuals(resid_model))))

# Recompute within/between decomposition on residualised predictions
# count_timepoints recomputed after confound filter so the 2+ filter is accurate
analysis_data_resid <- analysis_data_resid %>%
  group_by(participant_id) %>%
  mutate(
    stacked_pred_z_resid_mean = mean(stacked_pred_z_resid, na.rm = TRUE),
    count_timepoints_resid    = n()
  ) %>%
  ungroup() %>%
  mutate(
    stacked_pred_z_resid_dev = stacked_pred_z_resid - stacked_pred_z_resid_mean
  )

# Longitudinal subset for LME analyses
analysis_data_resid_lme <- analysis_data_resid %>%
  filter(count_timepoints_resid >= 2)

cat("Residualised longitudinal subset:",
    nrow(analysis_data_resid_lme), "observations from",
    n_distinct(analysis_data_resid_lme$participant_id), "participants.\n")

# Overall correlation between residualised predictions and observed g
r_resid <- cor(analysis_data_resid_lme$stacked_pred_z_resid,
               analysis_data_resid_lme$true_g_z,
               use = "complete.obs")
cat(sprintf("Pearson r (residualised pred vs true g): %.3f\n", r_resid))
```

# Residualised Correlations: Overall, Inter- and Intraindividual
```{r residualised_correlations}
# Interindividual: correlation between person-level means
participant_level_resid <- analysis_data_resid_lme %>%
  distinct(participant_id, .keep_all = TRUE)

r_resid_means <- cor(participant_level_resid$stacked_pred_z_resid_mean,
                     participant_level_resid$true_g_z_mean,
                     use = "complete.obs")

# Intraindividual: correlation between within-person deviations
r_resid_dev <- cor(analysis_data_resid_lme$stacked_pred_z_resid_dev,
                   analysis_data_resid_lme$true_g_z_dev,
                   use = "complete.obs")

cat("=== Residualised Prediction Correlations ===\n")
cat(sprintf("Overall:            r = %.3f,  R2 = %.4f  (N timepoints = %d)\n",
            r_resid, r_resid^2, nrow(analysis_data_resid_lme)))
cat(sprintf("Interindividual:    r = %.3f,  R2 = %.4f  (N participants = %d)\n",
            r_resid_means, r_resid_means^2, nrow(participant_level_resid)))
cat(sprintf("Intraindividual:    r = %.3f,  R2 = %.4f  (N timepoints = %d)\n",
            r_resid_dev, r_resid_dev^2, nrow(analysis_data_resid_lme)))
```

# Model 1: ADHD Additive (Residualised)
```{r m1_adhd_additive_resid}
m1_adhd_add_resid <- lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             ADHDdiagnosis + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m1_adhd_add_resid)
```

# Model 2: ADHD Interactive (Residualised)
```{r m2_adhd_interactive_resid}
m2_adhd_int_resid <- lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             ADHDdiagnosis +
             stacked_pred_z_resid_dev:ADHDdiagnosis +
             stacked_pred_z_resid_mean:ADHDdiagnosis +
             (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m2_adhd_int_resid)
```

# ADHD Model Comparison (Residualised)
```{r adhd_comparison_resid}
comp_adhd_resid <- anova(m1_adhd_add_resid, m2_adhd_int_resid)
print(comp_adhd_resid)

delta_bic_resid <- BIC(m2_adhd_int_resid) - BIC(m1_adhd_add_resid)
cat("\nDelta BIC (Interactive - Additive):", round(delta_bic_resid, 2), "\n")
if (delta_bic_resid > 0) {
  cat("Evidence favours the simpler ADDITIVE model.\n")
} else {
  cat("Evidence favours the INTERACTIVE model.\n")
}
```

# Model 3: Neuroimaging-only (Residualised)
```{r m3_neuroimaging_resid}
m3_neuro_resid <- lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m3_neuro_resid)
```

# Model 4: Age-only
```{r m4_age_resid}
# Age-only model: no residualisation needed (predictions not involved)
m4_age_resid <- lmer(
  true_g_z ~ age_deviation + age_mean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m4_age_resid)
```

# Model 5: Neuroimaging-and-Age (Residualised)
```{r m5_neuro_age_resid}
m5_neuro_age_resid <- lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             age_deviation + age_mean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m5_neuro_age_resid)
```

# Core Models Comparison (Residualised)
```{r core_comparison_resid}
comp_core_resid <- anova(m3_neuro_resid, m4_age_resid, m5_neuro_age_resid)
print(comp_core_resid)
```

# SJPlot Tables: Core Models (Residualised)
```{r sjplot_core_resid}
tab_model(m3_neuro_resid, m4_age_resid, m5_neuro_age_resid,
          dv.labels = c("Neuroimaging Only", "Age Only", "Neuroimaging-and-Age"),
          pred.labels = c(
            "(Intercept)"                  = "Intercept",
            "stacked_pred_z_resid_dev"     = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_resid_mean"    = "ĝ<sub>mean</sub>",
            "age_deviation"                = "age<sub>deviation</sub>",
            "age_mean"                     = "age<sub>mean</sub>"
          ),
          show.aic = TRUE, show.dev = TRUE,
          title = "Model Comparison: Neuroimaging and Age (Residualised)",
          file = "core_models_resid")
```

# SJPlot Tables: ADHD Models (Residualised)
```{r sjplot_adhd_resid}
tab_model(m1_adhd_add_resid, m2_adhd_int_resid,
          dv.labels = c("ADHD Additive", "ADHD Interactive"),
          pred.labels = c(
            "(Intercept)"                                 = "(Intercept)",
            "stacked_pred_z_resid_dev"                    = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_resid_mean"                   = "ĝ<sub>mean</sub>",
            "ADHDdiagnosisADHD"                           = "ADHD Diagnosis",
            "stacked_pred_z_resid_dev:ADHDdiagnosisADHD" = "ĝ<sub>deviation</sub> × ADHD",
            "stacked_pred_z_resid_mean:ADHDdiagnosisADHD" = "ĝ<sub>mean</sub> × ADHD"
          ),
          show.aic = TRUE, show.dev = TRUE,
          title = "ADHD Diagnosis Comparison (Residualised)",
          file = "adhd_diagnosis_resid")
```

# Model 6: Hyperactivity (Residualised)
```{r m6_hyperactivity_resid}
m6_hyper_resid <- lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             HyperactivityDeviation + HyperactivityMean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m6_hyper_resid)
```

# Hyperactivity Only (Residualised)
```{r m6p5_hyperactivity_only_resid}
m6p5_hyper_resid <- lmer(
  true_g_z ~ HyperactivityDeviation + HyperactivityMean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m6p5_hyper_resid)
```

# Model 7: Inattention (Residualised)
```{r m7_inattention_resid}
m7_inatt_resid <- lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             InattentionDeviation + InattentionMean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m7_inatt_resid)
```

# Inattention Only (Residualised)
```{r m7p5_inattention_only_resid}
m7p5_inatt_resid <- lmer(
  true_g_z ~ InattentionDeviation + InattentionMean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)
summary(m7p5_inatt_resid)
```

# SJPlot Tables: Symptom Models (Residualised)
```{r sjplot_symptoms_resid}
tab_model(m6p5_hyper_resid, m6_hyper_resid,
          dv.labels = c("Hyperactivity Only", "Neuroimaging-and-Hyperactivity"),
          pred.labels = c(
            "(Intercept)"                  = "Intercept",
            "stacked_pred_z_resid_dev"     = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_resid_mean"    = "ĝ<sub>mean</sub>",
            "HyperactivityMean"            = "Hyperactivity<sub>mean</sub>",
            "HyperactivityDeviation"       = "Hyperactivity<sub>deviation</sub>"
          ),
          show.aic = TRUE, show.dev = TRUE,
          title = "Hyperactivity (Residualised)",
          file = "hyperactivity_resid")

tab_model(m7p5_inatt_resid, m7_inatt_resid,
          dv.labels = c("Inattention Only", "Neuroimaging-and-Inattention"),
          pred.labels = c(
            "(Intercept)"                  = "Intercept",
            "stacked_pred_z_resid_dev"     = "ĝ<sub>deviation</sub>",
            "stacked_pred_z_resid_mean"    = "ĝ<sub>mean</sub>",
            "InattentionMean"              = "Inattention<sub>mean</sub>",
            "InattentionDeviation"         = "Inattention<sub>deviation</sub>"
          ),
          show.aic = TRUE, show.dev = TRUE,
          title = "Inattention (Residualised)",
          file = "inattention_resid")
```

# R2 Decomposition: Model 3 (Residualised)
```{r m3_r2_resid}
r2_m3_resid <- r2mlm::r2mlm(m3_neuro_resid)
print("Model 3 R-squared (residualised):")
print(r2_m3_resid)
```

# Export R2 Decomposition: Model 3 (Residualised)
```{r export_m3_r2_resid}
decomp_matrix_resid <- r2_m3_resid$Decompositions

decomp_df_resid <- data.frame(
  component = rownames(decomp_matrix_resid),
  total     = decomp_matrix_resid[, "total"],
  within    = decomp_matrix_resid[, "within"],
  between   = decomp_matrix_resid[, "between"],
  stringsAsFactors = FALSE
)

write.csv(decomp_df_resid, "m3_r2mlm_decomposition_resid.csv", row.names = FALSE)
cat("Exported to m3_r2mlm_decomposition_resid.csv\n")
```

# Commonality Analysis: Model 5 (Residualised)
```{r m5_commonality_resid}
m5_lme4_resid <- lme4::lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             age_deviation + age_mean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)

m5_commonality_results_resid <- glmm.hp::glmm.hp(m5_lme4_resid, type = "adjR2",
                                                   commonality = TRUE)
print("Model 5 Commonality (residualised):")
print(m5_commonality_results_resid)
```

# Adjust Commonality: Model 5 (Residualised)
```{r m5_adjust_commonality_resid}
m5_commonality_raw_resid <- m5_commonality_results_resid$commonality.analysis

is_total <- grepl("total|Total|TOTAL", rownames(m5_commonality_raw_resid), ignore.case = TRUE)
m5_components_resid <- m5_commonality_raw_resid[!is_total, , drop = FALSE]

m5_pos_resid  <- ifelse(m5_components_resid[, 1] < 0, 0, m5_components_resid[, 1])
names(m5_pos_resid) <- rownames(m5_components_resid)
m5_sum_resid  <- sum(m5_pos_resid)
m5_pct_resid  <- (m5_pos_resid / m5_sum_resid) * 100

m5_commonality_adjusted_resid <- data.frame(
  Component          = rownames(m5_components_resid),
  Original_Fraction  = m5_components_resid[, 1],
  Original_Percent   = m5_components_resid[, 2],
  Adjusted_Fraction  = m5_pos_resid,
  Adjusted_Percent   = m5_pct_resid,
  stringsAsFactors   = FALSE
)

m5_commonality_adjusted_resid <- rbind(
  m5_commonality_adjusted_resid,
  data.frame(Component = "Total",
             Original_Fraction = sum(m5_components_resid[, 1]),
             Original_Percent  = 100.00,
             Adjusted_Fraction = m5_sum_resid,
             Adjusted_Percent  = 100.00,
             stringsAsFactors  = FALSE)
)

print("Model 5 Adjusted Commonality (residualised):")
print(m5_commonality_adjusted_resid)
```

# Commonality Analysis: Model 6 Hyperactivity (Residualised)
```{r m6_commonality_resid}
m6_lme4_resid <- lme4::lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             HyperactivityDeviation + HyperactivityMean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)

m6_commonality_results_resid <- glmm.hp::glmm.hp(m6_lme4_resid, type = "adjR2",
                                                   commonality = TRUE)
print("Model 6 Commonality (residualised):")
print(m6_commonality_results_resid)
```

# Adjust Commonality: Model 6 (Residualised)
```{r m6_adjust_commonality_resid}
m6_commonality_raw_resid <- m6_commonality_results_resid$commonality.analysis

is_total <- grepl("total|Total|TOTAL", rownames(m6_commonality_raw_resid), ignore.case = TRUE)
m6_components_resid <- m6_commonality_raw_resid[!is_total, , drop = FALSE]

m6_pos_resid  <- ifelse(m6_components_resid[, 1] < 0, 0, m6_components_resid[, 1])
names(m6_pos_resid) <- rownames(m6_components_resid)
m6_sum_resid  <- sum(m6_pos_resid)
m6_pct_resid  <- (m6_pos_resid / m6_sum_resid) * 100

m6_commonality_adjusted_resid <- data.frame(
  Component          = rownames(m6_components_resid),
  Original_Fraction  = m6_components_resid[, 1],
  Original_Percent   = m6_components_resid[, 2],
  Adjusted_Fraction  = m6_pos_resid,
  Adjusted_Percent   = m6_pct_resid,
  stringsAsFactors   = FALSE
)

m6_commonality_adjusted_resid <- rbind(
  m6_commonality_adjusted_resid,
  data.frame(Component = "Total",
             Original_Fraction = sum(m6_components_resid[, 1]),
             Original_Percent  = 100.00,
             Adjusted_Fraction = m6_sum_resid,
             Adjusted_Percent  = 100.00,
             stringsAsFactors  = FALSE)
)

print("Model 6 Adjusted Commonality (residualised):")
print(m6_commonality_adjusted_resid)
```


# Commonality Analysis: Model 7 Inattention (Residualised)
```{r m7_commonality_resid}
m7_lme4_resid <- lme4::lmer(
  true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
             InattentionDeviation + InattentionMean + (1 | participant_id),
  data = analysis_data_resid_lme, REML = FALSE)

m7_commonality_results_resid <- glmm.hp::glmm.hp(m7_lme4_resid, type = "adjR2",
                                                   commonality = TRUE)
print("Model 7 Commonality (residualised):")
print(m7_commonality_results_resid)
```

# Adjust Commonality: Model 7 (Residualised)
```{r m7_adjust_commonality_resid}
m7_commonality_raw_resid <- m7_commonality_results_resid$commonality.analysis

is_total <- grepl("total|Total|TOTAL", rownames(m7_commonality_raw_resid), ignore.case = TRUE)
m7_components_resid <- m7_commonality_raw_resid[!is_total, , drop = FALSE]

m7_pos_resid  <- ifelse(m7_components_resid[, 1] < 0, 0, m7_components_resid[, 1])
names(m7_pos_resid) <- rownames(m7_components_resid)
m7_sum_resid  <- sum(m7_pos_resid)
m7_pct_resid  <- (m7_pos_resid / m7_sum_resid) * 100

m7_commonality_adjusted_resid <- data.frame(
  Component          = rownames(m7_components_resid),
  Original_Fraction  = m7_components_resid[, 1],
  Original_Percent   = m7_components_resid[, 2],
  Adjusted_Fraction  = m7_pos_resid,
  Adjusted_Percent   = m7_pct_resid,
  stringsAsFactors   = FALSE
)

m7_commonality_adjusted_resid <- rbind(
  m7_commonality_adjusted_resid,
  data.frame(Component = "Total",
             Original_Fraction = sum(m7_components_resid[, 1]),
             Original_Percent  = 100.00,
             Adjusted_Fraction = m7_sum_resid,
             Adjusted_Percent  = 100.00,
             stringsAsFactors  = FALSE)
)

print("Model 7 Adjusted Commonality (residualised):")
print(m7_commonality_adjusted_resid)
```

# R2 Decomposition: Model 6 Hyperactivity (Residualised)
```{r m6_r2_resid}
r2_m6_resid <- r2mlm::r2mlm(m6_lme4_resid)
print("Model 6 R-squared (residualised):")
print(r2_m6_resid)
```

# R2 Decomposition: Model 7 Inattention (Residualised)
```{r m7_r2_resid}
r2_m7_resid <- r2mlm::r2mlm(m7_lme4_resid)
print("Model 7 R-squared (residualised):")
print(r2_m7_resid)
```

# Export Commonality and Decomposition
```{r export_resid}
write.csv(decomp_df_resid,
          "m3_r2mlm_decomposition_resid.csv", row.names = FALSE)

write.csv(m5_commonality_adjusted_resid,
          "m5_neuro_age_commonality_adjusted_resid.csv", row.names = FALSE)

write.csv(m6_commonality_adjusted_resid,
          "m6_hyperactivity_commonality_adjusted_resid.csv", row.names = FALSE)

write.csv(m7_commonality_adjusted_resid,
          "m7_inattention_commonality_adjusted_resid.csv", row.names = FALSE)

cat("All residualised outputs exported.\n")
```

# Residualised Analysis Wrapper Function
```{r residualised_analysis_function}
# Wrapper: residualises stacked_pred_z from specified confounds, runs the full
# LME pipeline (m3, m5, m6, m7), and returns key comparison metrics.
# Uses the same logic as the main residualisation and analysis chunks above.
# get_adj_frac defined internally for self-containment.

run_residualised_analysis <- function(data_full, confound_vars, label) {

  # -- Internal helper: extract adjusted fraction by pattern match --
  .get_adj_frac <- function(df, pattern) {
    rows <- df[grepl(pattern, df$Component, ignore.case = FALSE, fixed = FALSE), ]
    rows <- rows[rows$Component != "Total", ]
    sum(rows$Adjusted_Fraction, na.rm = TRUE)
  }

  # -- Filter to complete cases for specified confounds --
  resid_data <- data_full
  for (v in confound_vars) {
    resid_data <- resid_data %>% filter(!is.na(.data[[v]]))
  }

  # -- OLS residualisation of stacked predictions --
  f_resid    <- as.formula(paste("stacked_pred_z ~", paste(confound_vars, collapse = " + ")))
  m_resid    <- lm(f_resid, data = resid_data)
  resid_data <- resid_data %>%
    mutate(stacked_pred_z_resid = as.double(scale(residuals(m_resid))))

  # -- Recompute within/between decomposition on residualised predictions --
  resid_data <- resid_data %>%
    group_by(participant_id) %>%
    mutate(
      stacked_pred_z_resid_mean = mean(stacked_pred_z_resid, na.rm = TRUE),
      count_timepoints_resid    = n()
    ) %>%
    ungroup() %>%
    mutate(stacked_pred_z_resid_dev = stacked_pred_z_resid - stacked_pred_z_resid_mean)

  # -- Overall r on full dataset (pre-2+ filter, apples-to-apples with r=0.459) --
  r_overall <- cor(resid_data$stacked_pred_z_resid, resid_data$true_g_z,
                   use = "complete.obs")

  # -- Longitudinal subset for LME analyses --
  lme_data <- resid_data %>% filter(count_timepoints_resid >= 2)

  # -- Interindividual and intraindividual correlations --
  pl      <- lme_data %>% distinct(participant_id, .keep_all = TRUE)
  r_means <- cor(pl$stacked_pred_z_resid_mean, pl$true_g_z_mean, use = "complete.obs")
  r_dev   <- cor(lme_data$stacked_pred_z_resid_dev, lme_data$true_g_z_dev,
                 use = "complete.obs")

  # -- Model 3: neuroimaging only --
  m3     <- lmer(true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
                   (1 | participant_id),
                 data = lme_data, REML = FALSE)
  r2_m3  <- r2mlm::r2mlm(m3)
  r2_intra <- r2_m3$R2s["f1", "within"]  * 100
  r2_inter <- r2_m3$R2s["f2", "between"] * 100

  # -- Commonality helper: adjust, zero negatives, return data frame --
  .run_commonality <- function(model) {
    comm_res <- glmm.hp::glmm.hp(model, type = "adjR2", commonality = TRUE)
    raw      <- comm_res$commonality.analysis
    is_tot   <- grepl("total|Total|TOTAL", rownames(raw), ignore.case = TRUE)
    comp     <- raw[!is_tot, , drop = FALSE]
    pos      <- ifelse(comp[, 1] < 0, 0, comp[, 1])
    names(pos) <- rownames(comp)
    data.frame(Component = rownames(comp), Adjusted_Fraction = pos,
               stringsAsFactors = FALSE)
  }

  # -- Model 5: neuroimaging + age (age capture percentages) --
  m5    <- lme4::lmer(true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
                        age_deviation + age_mean + (1 | participant_id),
                      data = lme_data, REML = FALSE)
  m5_df <- .run_commonality(m5)

  shared_dev_age  <- .get_adj_frac(m5_df, "resid_dev, and age_deviation")
  unique_age_dev  <- .get_adj_frac(m5_df, "Unique to age_deviation")
  age_intra       <- shared_dev_age  / (shared_dev_age  + unique_age_dev)  * 100

  shared_mean_age <- .get_adj_frac(m5_df, "resid_mean, and age_mean")
  unique_age_mean <- .get_adj_frac(m5_df, "Unique to age_mean")
  age_inter       <- shared_mean_age / (shared_mean_age + unique_age_mean) * 100

  # -- Model 6: hyperactivity commonality --
  m6    <- lme4::lmer(true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
                        HyperactivityDeviation + HyperactivityMean + (1 | participant_id),
                      data = lme_data, REML = FALSE)
  m6_df <- .run_commonality(m6)

  neuro_hyp_mean <- .get_adj_frac(m6_df, "HyperactivityMean") -
                    .get_adj_frac(m6_df, "Unique to HyperactivityMean")
  total_hyp_mean <- .get_adj_frac(m6_df, "HyperactivityMean")
  neuro_hyp_dev  <- .get_adj_frac(m6_df, "HyperactivityDeviation") -
                    .get_adj_frac(m6_df, "Unique to HyperactivityDeviation")
  total_hyp_dev  <- .get_adj_frac(m6_df, "HyperactivityDeviation")

  hyp_inter <- neuro_hyp_mean / total_hyp_mean * 100
  hyp_intra <- if (total_hyp_dev == 0) 100 else neuro_hyp_dev / total_hyp_dev * 100
  hyp_total <- (neuro_hyp_mean + neuro_hyp_dev) / (total_hyp_mean + total_hyp_dev) * 100

  # -- Model 7: inattention commonality --
  m7    <- lme4::lmer(true_g_z ~ stacked_pred_z_resid_dev + stacked_pred_z_resid_mean +
                        InattentionDeviation + InattentionMean + (1 | participant_id),
                      data = lme_data, REML = FALSE)
  m7_df <- .run_commonality(m7)

  neuro_inatt_mean <- .get_adj_frac(m7_df, "InattentionMean") -
                      .get_adj_frac(m7_df, "Unique to InattentionMean")
  total_inatt_mean <- .get_adj_frac(m7_df, "InattentionMean")
  inatt_inter      <- neuro_inatt_mean / total_inatt_mean * 100

  cat(sprintf("\n==== %s ====\n", label))
  cat(sprintf("N (complete confound cases): %d  |  N (LME subset): %d\n",
              nrow(resid_data), nrow(lme_data)))
  cat(sprintf("Overall r:    %.3f\n", r_overall))
  cat(sprintf("r means:      %.3f   r dev:   %.3f\n", r_means, r_dev))
  cat(sprintf("Intra R2:     %.2f%%  Inter R2: %.2f%%\n", r2_intra, r2_inter))
  cat(sprintf("Age intra:    %.1f%%  Age inter: %.1f%%\n", age_intra, age_inter))
  cat(sprintf("Hyp inter:    %.1f%%  Hyp intra: %.1f%%  Hyp total: %.1f%%\n",
              hyp_inter, hyp_intra, hyp_total))
  cat(sprintf("Inatt inter:  %.1f%%\n", inatt_inter))

  invisible(list(
    label       = label,
    r_overall   = r_overall,
    r_means     = r_means,
    r_dev       = r_dev,
    r2_intra    = r2_intra,
    r2_inter    = r2_inter,
    age_intra   = age_intra,
    age_inter   = age_inter,
    hyp_inter   = hyp_inter,
    hyp_intra   = hyp_intra,
    hyp_total   = hyp_total,
    inatt_inter = inatt_inter
  ))
}
```

# Separate Residualisations: Single Confounds
```{r separate_residualisations}
# Runs the full analysis pipeline separately for each confound in isolation,
# plus the combined residualisation, for side-by-side comparison in the table.
# NOTE: the combined residualisation here uses the wrapper function for
# consistency; the detailed analysis above remains for full output/exports.

cat("Running single-confound residualisations. This may take a few minutes.\n")

res_med     <- run_residualised_analysis(
  analysis_data_full, "current_stimmed_bin",             "Medication only")

res_volumes <- run_residualised_analysis(
  analysis_data_full, "perc_remaining_post_censoring",   "Volumes remaining only")

res_edu     <- run_residualised_analysis(
  analysis_data_full, "highest_edu",                     "Parental education only")

res_euler   <- run_residualised_analysis(
  analysis_data_full, "mean_euler",                      "Euler number only")

res_combined <- run_residualised_analysis(
  analysis_data_full,
  c("current_stimmed_bin", "perc_remaining_post_censoring", "highest_edu", "mean_euler"),
  "Combined (all four)")

cat("\nAll separate residualisations complete.\n")
```

# Summary Comparison Table
```{r summary_comparison_table}
# ── Correlations (residualised) ──────────────────────────────────────────────
# r_resid, r_resid_means, r_resid_dev computed in residualised_correlations chunk

# ── R2 decomposition ─────────────────────────────────────────────────────────
r2_intra_resid <- r2_m3_resid$R2s["f1", "within"]   * 100
r2_inter_resid <- r2_m3_resid$R2s["f2", "between"]  * 100

# ── Helper: extract Adjusted_Fraction by partial component name match ────────
get_adj_frac <- function(df, pattern) {
  rows <- df[grepl(pattern, df$Component, ignore.case = FALSE, fixed = FALSE), ]
  rows <- rows[rows$Component != "Total", ]
  sum(rows$Adjusted_Fraction, na.rm = TRUE)
}

# ── Age capture percentages (from m5 commonality) ────────────────────────────
shared_dev_age_dev    <- get_adj_frac(m5_commonality_adjusted_resid,
                                      "resid_dev, and age_deviation")
unique_age_dev        <- get_adj_frac(m5_commonality_adjusted_resid,
                                      "Unique to age_deviation")
age_intra_resid       <- shared_dev_age_dev / (shared_dev_age_dev + unique_age_dev) * 100

shared_mean_age_mean  <- get_adj_frac(m5_commonality_adjusted_resid,
                                      "resid_mean, and age_mean")
unique_age_mean       <- get_adj_frac(m5_commonality_adjusted_resid,
                                      "Unique to age_mean")
age_inter_resid       <- shared_mean_age_mean / (shared_mean_age_mean + unique_age_mean) * 100

# ── Hyperactivity capture percentages (from m6 commonality) ──────────────────
# All terms involving neuroimaging AND HypMean (interindividual)
neuro_hyp_mean_resid  <- get_adj_frac(m6_commonality_adjusted_resid, "HyperactivityMean") -
                         get_adj_frac(m6_commonality_adjusted_resid, "Unique to HyperactivityMean")
total_hyp_mean_resid  <- get_adj_frac(m6_commonality_adjusted_resid, "HyperactivityMean")
hyp_inter_resid       <- neuro_hyp_mean_resid / total_hyp_mean_resid * 100

# All terms involving neuroimaging AND HypDev (intraindividual)
neuro_hyp_dev_resid   <- get_adj_frac(m6_commonality_adjusted_resid, "HyperactivityDeviation") -
                         get_adj_frac(m6_commonality_adjusted_resid, "Unique to HyperactivityDeviation")
total_hyp_dev_resid   <- get_adj_frac(m6_commonality_adjusted_resid, "HyperactivityDeviation")
hyp_intra_resid       <- if (total_hyp_dev_resid == 0) 100 else
                          neuro_hyp_dev_resid / total_hyp_dev_resid * 100

# Total hyperactivity capture
total_hyp_resid       <- total_hyp_mean_resid + total_hyp_dev_resid
neuro_hyp_total_resid <- neuro_hyp_mean_resid + neuro_hyp_dev_resid
hyp_total_resid       <- neuro_hyp_total_resid / total_hyp_resid * 100

# ── Inattention capture percentages (from m7 commonality) ────────────────────
neuro_inatt_mean_resid <- get_adj_frac(m7_commonality_adjusted_resid, "InattentionMean") -
                          get_adj_frac(m7_commonality_adjusted_resid, "Unique to InattentionMean")
total_inatt_mean_resid <- get_adj_frac(m7_commonality_adjusted_resid, "InattentionMean")
inatt_inter_resid      <- neuro_inatt_mean_resid / total_inatt_mean_resid * 100

# ── Original values (from primary analysis script) ───────────────────────────
# r values from Figure 5 scatterplots and manuscript text
orig <- list(
  r_overall   = 0.459,
  r_means     = 0.47,
  r_dev       = 0.52,
  r2_intra    = 18.82,
  r2_inter    = 25.01,
  age_intra   = 60.87,
  age_inter   = 72.52,
  hyp_inter   = 43.36,
  hyp_intra   = 100.00,
  hyp_total   = 58.79,
  inatt_inter = 25.99
)

# ── Build table ───────────────────────────────────────────────────────────────
comparison_table <- data.frame(
  Metric = c(
    "Overall r (predicted vs observed g)",
    "r (interindividual: means)",
    "r (intraindividual: deviations)",
    "Intraindividual R2 (f1 within)",
    "Interindividual R2 (f2 between)",
    "Intraindividual age-related change captured (%)",
    "Interindividual age-related change captured (%)",
    "Hyperactivity-cognition link captured: interindividual (%)",
    "Hyperactivity-cognition link captured: intraindividual (%)",
    "Hyperactivity-cognition link captured: total (%)",
    "Inattention-cognition link captured: interindividual (%)"
  ),
  Original = round(c(
    orig$r_overall, orig$r_means, orig$r_dev,
    orig$r2_intra, orig$r2_inter,
    orig$age_intra, orig$age_inter,
    orig$hyp_inter, orig$hyp_intra, orig$hyp_total,
    orig$inatt_inter
  ), 2),
  Resid_medication = round(c(
    res_med$r_overall, res_med$r_means, res_med$r_dev,
    res_med$r2_intra, res_med$r2_inter,
    res_med$age_intra, res_med$age_inter,
    res_med$hyp_inter, res_med$hyp_intra, res_med$hyp_total,
    res_med$inatt_inter
  ), 2),
  Resid_volumes = round(c(
    res_volumes$r_overall, res_volumes$r_means, res_volumes$r_dev,
    res_volumes$r2_intra, res_volumes$r2_inter,
    res_volumes$age_intra, res_volumes$age_inter,
    res_volumes$hyp_inter, res_volumes$hyp_intra, res_volumes$hyp_total,
    res_volumes$inatt_inter
  ), 2),
  Resid_education = round(c(
    res_edu$r_overall, res_edu$r_means, res_edu$r_dev,
    res_edu$r2_intra, res_edu$r2_inter,
    res_edu$age_intra, res_edu$age_inter,
    res_edu$hyp_inter, res_edu$hyp_intra, res_edu$hyp_total,
    res_edu$inatt_inter
  ), 2),
  Resid_euler = round(c(
    res_euler$r_overall, res_euler$r_means, res_euler$r_dev,
    res_euler$r2_intra, res_euler$r2_inter,
    res_euler$age_intra, res_euler$age_inter,
    res_euler$hyp_inter, res_euler$hyp_intra, res_euler$hyp_total,
    res_euler$inatt_inter
  ), 2),
  Resid_combined = round(c(
    res_combined$r_overall, res_combined$r_means, res_combined$r_dev,
    res_combined$r2_intra, res_combined$r2_inter,
    res_combined$age_intra, res_combined$age_inter,
    res_combined$hyp_inter, res_combined$hyp_intra, res_combined$hyp_total,
    res_combined$inatt_inter
  ), 2),
  stringsAsFactors = FALSE
)

print(comparison_table)
write.csv(comparison_table, "sensitivity_residualised_comparison_table.csv", row.names = FALSE)
cat("\nExported to sensitivity_residualised_comparison_table.csv\n")
```
