---
title: "Regularized G-Factor Model for OREGON1000 Cognitive Data"
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

This R Notebook implements a regularized g-factor model for cognitive data from the OREGON1000 dataset. The regularized approach adds a small constant to the diagonal of the covariance matrix to improve numerical stability while maintaining all variables as continuous.

# Libraries

```{r load_libraries}
library(data.table)
library(psych) 
library(lavaan)
library(lavaanPlot)
library(ggplot2)
library(corrplot)
library(dplyr)
```

# Set working directory

```{r set_directory}
# Set working directory
knitr::opts_knit$set(root.dir ='/media/hcs-sci-psy-narun/Jack/g_factor_analysis_v5_nested')
```

# Upload train and test sets 

```{r load_data}
base_path <- "/media/hcs-sci-psy-narun/Jack/g_factor_analysis_v5_nested/fold_"
file_names <- c("cogscores_12T_i2_train_with_id_scaled_fold_", "cogscores_12T_i2_test_with_id_scaled_fold_")

for (i in 0:4) {
  assign(paste0("train_", i), read.csv(paste0(base_path, i, "/cogscores/", file_names[1], i, ".csv"), header = TRUE))
  assign(paste0("test_", i), read.csv(paste0(base_path, i, "/cogscores/", file_names[2], i, ".csv"), header = TRUE))
}
```

# Validate independent fold structure

```{r validate_folds}
# Function to extract subject IDs from timepoint IDs
extract_subject_id <- function(timepoint_ids) {
  subject_ids <- sapply(strsplit(timepoint_ids, "_"), function(x) x[1])
  return(unique(subject_ids))
}

# Updated validation function for nested cross-validation
validate_nested_cv_folds <- function() {
  cat("=== NESTED CV FOLD VALIDATION ===\n")
  validation_passed <- TRUE
  
  # Track subject appearances across all folds
  subject_appearances <- list()
  
  # Get all subjects from all folds and track their appearances
  for (i in 0:4) {
    # Get train and test dataframes
    train_df <- get(paste0("train_", i))
    test_df <- get(paste0("test_", i))
    
    # Extract subject IDs from timepoint IDs
    train_subjects <- extract_subject_id(train_df$eid)
    test_subjects <- extract_subject_id(test_df$eid)
    
    cat(paste("Outer Fold", i, ": Train subjects =", length(train_subjects), 
              ", Test subjects =", length(test_subjects), "\n"))
    
    # Check for subject overlap between train and test WITHIN this outer fold
    overlap <- intersect(train_subjects, test_subjects)
    if (length(overlap) > 0) {
      cat(paste("VALIDATION FAILED! Outer fold", i, "has", length(overlap), 
                "subjects in both train and test sets!\n"))
      validation_passed <- FALSE
    } else {
      cat(paste("✓ Outer fold", i, ": No subject overlap between train and test\n"))
    }
    
    # Track appearances for nested CV validation
    for (subject in train_subjects) {
      if (is.null(subject_appearances[[subject]])) {
        subject_appearances[[subject]] <- list(train_folds = c(), test_folds = c())
      }
      subject_appearances[[subject]]$train_folds <- c(subject_appearances[[subject]]$train_folds, i)
    }
    
    for (subject in test_subjects) {
      if (is.null(subject_appearances[[subject]])) {
        subject_appearances[[subject]] <- list(train_folds = c(), test_folds = c())
      }
      subject_appearances[[subject]]$test_folds <- c(subject_appearances[[subject]]$test_folds, i)
    }
  }
  
  # Validate nested CV pattern: each subject should appear in exactly 4 train folds and 1 test fold
  nested_cv_valid <- TRUE
  for (subject in names(subject_appearances)) {
    n_train <- length(subject_appearances[[subject]]$train_folds)
    n_test <- length(subject_appearances[[subject]]$test_folds)
    
    if (n_train != 4 || n_test != 1) {
      cat(paste("NESTED CV VIOLATION! Subject", subject, "appears in", n_train, 
                "train folds and", n_test, "test folds (should be 4 train, 1 test)\n"))
      nested_cv_valid <- FALSE
      validation_passed <- FALSE
    }
  }
  
  if (nested_cv_valid) {
    cat("✓ Nested CV structure is valid: Each subject appears in exactly 4 train folds and 1 test fold\n")
  }
  
  # Final validation status
  if (validation_passed) {
    cat("✓ All nested CV validations passed\n")
  } else {
    cat("✗ VALIDATION FAILED: Issues detected in nested CV structure\n")
  }
  cat("=== NESTED CV VALIDATION COMPLETE ===\n")
  
  return(validation_passed)
}

# Run the updated validation
fold_validation_result <- validate_nested_cv_folds()
```

# Parse timepoint IDs and prepare data

```{r parse_timepoints}
# Function to extract components from timepoint ID
extract_timepoint_info <- function(df) {
  # Create a copy to avoid modifying the original
  result <- df
  
  # Extract subject ID and age information
  result$subject_id <- sapply(result$eid, function(id) {
    parts <- strsplit(as.character(id), "_")[[1]]
    return(parts[1])
  })
  
  # Extract age information if available
  result$age_info <- sapply(result$eid, function(id) {
    parts <- strsplit(as.character(id), "_")[[1]]
    if (length(parts) > 1) {
      age_part <- parts[2]
      if (grepl("^Age", age_part)) {
        return(age_part)
      }
    }
    return(NA)
  })
  
  # Try to extract numeric age in years
  result$age_years <- sapply(result$age_info, function(age_info) {
    if (is.na(age_info)) return(NA)
    if (grepl("^Age[0-9]{2}[0-9]$", age_info)) {
      # Format is "AgeXXY" where XX=years, Y=months
      years <- as.numeric(substr(age_info, 4, 5))
      months <- as.numeric(substr(age_info, 6, 6))
      return(years + months/12)
    }
    return(NA)
  })
  
  return(result)
}

# Apply the function to each dataset
for (i in 0:4) {
  train_name <- paste0("train_", i)
  test_name <- paste0("test_", i)
  
  assign(train_name, extract_timepoint_info(get(train_name)))
  assign(test_name, extract_timepoint_info(get(test_name)))
}
```


# Create nice variable names for visualization

```{r nice_names}
# Create mapping for nice display names
nice_names <- c(
  # Stop Signal Task
  stop_sdgort = "Stop Task SD of Go RT",
  
  # Continuous Performance Task
  dprime1 = "CPT D-prime (Catch)",
  
  # N-Back Task
  n2backacc = "N-Back 2-back Accuracy",
  
  # WISC Digit Span
  digitspan_dsfscore = "Digit Span Forward",
  digitspan_dsbscore = "Digit Span Backward",
  
  # WIAT Achievement Test
  wordread_score = "WIAT Word Reading"
)

# Helper function to get nice names
get_nice_name <- function(var_name) {
  for (pattern in names(nice_names)) {
    if (grepl(pattern, var_name, fixed = TRUE)) {
      return(nice_names[pattern])
    }
  }
  return(var_name)
}
```

# Remove non-cognitive columns for analysis

```{r remove_non_cognitive}
# Function to remove non-cognitive columns
remove_columns <- function(df) {
  # Keep a copy of the eid column to add back later
  eid_col <- df$eid
  
  # Remove non-cognitive columns
  df <- subset(df, select = -c(eid, subject_id, age_info, age_years))
  
  return(list(df = df, eid = eid_col))
}

# Apply the function to each data frame
for (i in 0:4) {
  train_name <- paste0("train_", i)
  test_name <- paste0("test_", i)
  
  # Process train set
  train_result <- remove_columns(get(train_name))
  assign(paste0("train_", i, "_eid"), train_result$eid)
  assign(train_name, train_result$df)
  
  # Process test set
  test_result <- remove_columns(get(test_name))
  assign(paste0("test_", i, "_eid"), test_result$eid)
  assign(test_name, test_result$df)
}
```

```{r missingness}
# Check missingness patterns
cognitive_vars <- c("n2backacc", "dprime1", "stop_sdgort", 
                   "digitspan_dsfscore", "digitspan_dsbscore", "wordread_score")
sapply(cognitive_vars, function(x) sum(is.na(x))/length(x))

```


# Examine correlation patterns in dataset

```{r correlation_matrix, fig.width=10, fig.height=8}
# Calculate correlation matrix for fold 0
correlation_matrix <- cor(train_0, use = "pairwise.complete.obs")

# Create nicer labels for visualization
nice_labels <- sapply(colnames(correlation_matrix), get_nice_name)
rownames(correlation_matrix) <- nice_labels
colnames(correlation_matrix) <- nice_labels

# Create improved correlation plot
corrplot(correlation_matrix, 
         method = "color",       # Colored squares
         type = "upper",         # Only upper triangle 
         tl.col = "black",       # Black text labels
         tl.srt = 45,            # Angled text
         tl.cex = 0.9,           # Slightly larger text
         cl.cex = 0.9,           # Slightly larger legend text
         addCoef.col = "black",  # Black numbers
         number.cex = 0.8,       # Slightly smaller numbers
         number.digits = 2,      # Only show 2 decimal places
         diag = FALSE,           # Don't show diagonal (always 1)
         title = "Correlation Matrix of Cognitive Variables")
```

# Regularized G-Factor Model

## Create direct model syntax

```{r define_model}
# Create a direct one-factor model (g factor) syntax
direct_model_syntax <- "g =~ stopsig_stop_sdgort + cpth_dprime1 + nback_n2backacc + 
                         wisc_digitspan_dsfscore + wisc_digitspan_dsbscore + 
                         wiat_wordread_score"
```

## Function for regularized model fitting

```{r regularized_approach}
# Function to fit with regularized covariance
fit_regularized <- function(data, fold_idx) {
  cat(paste0("\n====== Fold ", fold_idx, " - Regularized Approach ======\n"))
  
  tryCatch({
    # Calculate covariance matrix
    cov_matrix <- cov(data, use = "pairwise.complete.obs")
    
    # Add small constant to diagonal for regularization
    diag(cov_matrix) <- diag(cov_matrix) + 0.05
    
    # Fit model using the regularized covariance matrix
    model_fit <- cfa(direct_model_syntax, 
                     sample.cov = cov_matrix,
                     sample.nobs = nrow(data),
                     std.lv = TRUE,
                     missing = "fiml")  # FIML for handling any remaining missing data
    
    fit_summary <- summary(model_fit, fit.measures = TRUE, standardized = TRUE)
    print(fit_summary)
    return(model_fit)
  }, error = function(e) {
    cat("Model did not converge. Error:", e$message, "\n")
    return(NULL)
  })
}
```

## Fit models for each fold

```{r fit_models}
# Fit models for each fold
regularized_fits <- list()
for (i in 0:4) {
  regularized_fits[[i+1]] <- fit_regularized(get(paste0("train_", i)), i)
}
```

## Visualize the model

```{r model_visualization, fig.width=10, fig.height=8}
# Function to create lavaan plot for the model
create_lavaan_plot <- function(model) {
  if (is.null(model)) return(NULL)
  
  # Create a model visualization
  plot <- lavaanPlot(model = model, 
                    node_options = list(shape = "box", fontname = "Helvetica"),
                    edge_options = list(color = "grey"), 
                    coefs = TRUE,
                    stand = TRUE,
                    sig = 0.05)
  
  return(plot)
}

# Create and show plot for first fold
model_plot <- create_lavaan_plot(regularized_fits[[1]])
if (!is.null(model_plot)) {
  print(model_plot)
}
```

## Examine loadings across folds

```{r examine_loadings, fig.width=10, fig.height=6}
# Function to extract loadings
extract_loadings <- function(model) {
  if (is.null(model)) return(NULL)
  
  tryCatch({
    # Get standardized solution
    solution <- standardizedSolution(model)
    
    # Extract loadings (rows with operator "=~")
    loadings <- solution[solution$op == "=~", ]
    
    # Create a data frame with variable names and loadings
    result <- data.frame(
      Variable = loadings$rhs,
      Loading = loadings$est.std
    )
    
    # Add nice names
    result$Nice_Name <- sapply(result$Variable, get_nice_name)
    
    return(result)
  }, error = function(e) {
    return(NULL)
  })
}

# Get and compare loadings across folds
all_loadings <- data.frame()

for (i in 0:4) {
  fold_loadings <- extract_loadings(regularized_fits[[i+1]])
  
  if (!is.null(fold_loadings)) {
    fold_loadings$Fold <- paste("Fold", i)
    all_loadings <- rbind(all_loadings, fold_loadings)
  }
}

# Create comparison plot
if (nrow(all_loadings) > 0) {
  ggplot(all_loadings, aes(x = Nice_Name, y = Loading, fill = Fold)) +
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip() +
    labs(title = "Comparison of G-Factor Loadings Across Folds",
         x = "Cognitive Variable",
         y = "Loading") +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```

# Extract G-Factor Scores

```{r extract_scores}
save_path <- '/media/hcs-sci-psy-narun/Jack/g_factor_analysis_v6_nested_rework/'

# Calculate g-factor scores using regression method
for (i in 0:4) {
  # Skip if model did not converge
  if (is.null(regularized_fits[[i+1]])) {
    cat(paste0("Skipping fold ", i, " as model did not converge\n"))
    next
  }
  
  # Get model
  model <- regularized_fits[[i+1]]
  
  # Get data
  train_data <- get(paste0("train_", i))
  test_data <- get(paste0("test_", i))
  
  # Get original IDs
  train_ids <- get(paste0("train_", i, "_eid"))
  test_ids <- get(paste0("test_", i, "_eid"))
  
  tryCatch({
    # Extract standardized loadings for g-factor
    loadings <- standardizedSolution(model)
    g_loadings <- loadings[loadings$op == "=~" & loadings$lhs == "g", ]
    
    # Variables in the model
    model_vars <- g_loadings$rhs
    
    # FIXED function to compute g-factor scores (handles pre-scaled data)
    compute_g_scores <- function(data, vars, loadings, fold_name = paste0("fold_", i)) {
      cat(paste0("\n--- Computing G-scores for ", fold_name, " ---\n"))
      
      # Check if data is already standardized
      means <- colMeans(data[, vars], na.rm = TRUE)
      sds <- apply(data[, vars], 2, sd, na.rm = TRUE)
      
      cat("Input data statistics:\n")
      for (j in 1:min(3, length(vars))) {
        cat(paste0("  ", vars[j], ": mean=", round(means[j], 4), ", sd=", round(sds[j], 4), "\n"))
      }
      
      # Check if data appears to be already standardized
      already_scaled <- all(abs(means) < 0.2) && all(abs(sds - 1) < 0.3)
      
      if (already_scaled) {
        cat("*** DATA APPEARS ALREADY STANDARDIZED - USING DIRECTLY ***\n")
        cat("*** AVOIDING DOUBLE-STANDARDIZATION ***\n")
        
        # Use data as-is since it's already scaled
        scaled_data <- as.matrix(data[, vars])
        
        # Replace NAs with 0 (which is the mean in standardized data)
        scaled_data[is.na(scaled_data)] <- 0
        
      } else {
        cat("*** DATA NOT STANDARDIZED - APPLYING STANDARDIZATION ***\n")
        
        # Create matrix for scaled data
        scaled_data <- matrix(NA, nrow = nrow(data), ncol = length(vars))
        colnames(scaled_data) <- vars
        
        # Scale each column
        for (j in 1:length(vars)) {
          var <- vars[j]
          if (sds[j] > 0) {
            scaled_data[, j] <- (data[, var] - means[j]) / sds[j]
          } else {
            scaled_data[, j] <- 0  # Handle zero variance
            cat(paste0("Warning: Zero variance in ", var, "\n"))
          }
        }
        
        # Replace NAs with 0 (mean after standardization)
        scaled_data[is.na(scaled_data)] <- 0
      }
      
      # Compute scores using loadings
      scores <- scaled_data %*% matrix(loadings$est.std, ncol = 1)
      
      cat("Pre-standardization g-scores:\n")
      cat(paste0("  Mean: ", round(mean(scores), 4), "\n"))
      cat(paste0("  SD: ", round(sd(scores), 4), "\n"))
      cat(paste0("  Range: [", round(min(scores), 4), ", ", round(max(scores), 4), "]\n"))
      
      # Standardize the final scores to have mean 0 and SD 1
      scores <- scale(scores)
      
      cat("Final g-scores:\n")
      cat(paste0("  Mean: ", round(mean(scores), 4), "\n"))
      cat(paste0("  SD: ", round(sd(scores), 4), "\n"))
      cat(paste0("  Range: [", round(min(scores), 4), ", ", round(max(scores), 4), "]\n"))
      
      # CRITICAL: Check correlations with original cognitive variables to detect inversion
      cat("G-score correlations with original cognitive variables:\n")
      inversion_detected <- FALSE
      negative_correlations <- 0
      
      for (var in vars) {
        if (var %in% names(data)) {
          orig_values <- data[[var]]
          valid_indices <- !is.na(orig_values)
          
          if (sum(valid_indices) > 10) {
            corr_val <- cor(scores[valid_indices], orig_values[valid_indices], use = "complete.obs")
            cat(paste0("  ", var, ": r=", round(corr_val, 4)))
            
            # Most cognitive tests should correlate positively with g-factor
            # Exception: stop_sdgort (RT variability) should be negative
            if (grepl("stop.*sdgort", var)) {
              # RT variability should be negative
              if (corr_val > -0.1) {
                cat(" *** EXPECTED NEGATIVE but got positive - possible issue ***")
              } else {
                cat(" ✓ (expected negative)")
              }
            } else {
              # All other tests should be positive
              if (corr_val < -0.1) {
                cat(" *** NEGATIVE - POTENTIAL PROBLEM ***")
                negative_correlations <- negative_correlations + 1
              } else {
                cat(" ✓ (expected positive)")
              }
            }
            cat("\n")
          }
        }
      }
      
      # Check for systematic inversion
      expected_positive_vars <- length(vars) - 1  # All except stop_sdgort
      if (negative_correlations > expected_positive_vars * 0.5) {
        cat("*** WARNING: MAJORITY OF CORRELATIONS ARE NEGATIVE ***\n")
        cat("*** THIS SUGGESTS G-FACTOR MAY BE INVERTED ***\n")
        inversion_detected <- TRUE
      }
      
      # If inversion detected, could apply correction here
      if (inversion_detected) {
        cat("*** APPLYING AUTOMATIC INVERSION CORRECTION ***\n")
        scores <- -scores
        scores <- scale(scores)  # Re-standardize after flip
        cat("G-factor scores have been inverted to correct direction\n")
      }
      
      return(scores)
    }
    
    # Compute scores with enhanced diagnostics
    g_train_scores <- compute_g_scores(train_data, model_vars, g_loadings, paste0("fold_", i, "_train"))
    g_test_scores <- compute_g_scores(test_data, model_vars, g_loadings, paste0("fold_", i, "_test"))
    
    # Create data frames with IDs
    g_train <- data.frame(g = as.numeric(g_train_scores), eid = train_ids)
    g_test <- data.frame(g = as.numeric(g_test_scores), eid = test_ids)
    
    # Save to global environment
    assign(paste0("g_train_", i), g_train, envir = .GlobalEnv)
    assign(paste0("g_test_", i), g_test, envir = .GlobalEnv)
    
    # Create directories if they don't exist
    dir.create(file.path(save_path, paste0("fold_", i), "g"), recursive = TRUE, showWarnings = FALSE)
    
    # Save as csv
    write.csv(g_train, file = file.path(save_path, paste0("fold_", i), "g", paste0("g_train_regularized_", i, ".csv")),
              row.names = FALSE)
    write.csv(g_test, file = file.path(save_path, paste0("fold_", i), "g", paste0("g_test_regularized_", i, ".csv")), row.names = FALSE)
    
    cat(paste0("Saved g-factor scores for fold ", i, "\n"))
  }, error = function(e) {
    cat(paste0("Error extracting g-factor scores for fold ", i, ": ", e$message, "\n"))
  })
}
```

# Combine all g-factor scores

```{r combine_scores}
# Updated score combination for nested CV
combine_nested_cv_scores <- function() {
  # Initialize tracking structures
  all_g_train <- data.frame()
  all_g_test <- data.frame()
  
  # Track which fold each score comes from
  for (i in 0:4) {
    # Check if g-factor scores exist for this fold
    if (exists(paste0("g_train_", i))) {
      fold_train <- get(paste0("g_train_", i))
      fold_train$outer_fold <- i
      fold_train$split_type <- "train"
      all_g_train <- rbind(all_g_train, fold_train)
    }
    
    if (exists(paste0("g_test_", i))) {
      fold_test <- get(paste0("g_test_", i))
      fold_test$outer_fold <- i
      fold_test$split_type <- "test"
      all_g_test <- rbind(all_g_test, fold_test)
    }
  }
  
  # Combine all scores with fold information
  all_g_scores <- rbind(all_g_train, all_g_test)
  
  # Extract subject IDs for analysis
  all_g_scores$subject_id <- sapply(all_g_scores$eid, function(id) {
    strsplit(as.character(id), "_")[[1]][1]
  })
  
  # Create separate datasets for different analyses
  
  # 1. Test scores only (each subject appears exactly once)
  test_scores_only <- all_g_scores[all_g_scores$split_type == "test", ]
  cat("Test scores (each subject once):", nrow(test_scores_only), "scores from", 
      length(unique(test_scores_only$subject_id)), "unique subjects\n")
  
  # 2. All scores (for comprehensive analysis)
  cat("All scores (subjects appear multiple times):", nrow(all_g_scores), "total scores\n")
  
  # 3. Average training scores per subject (average across the 4 training appearances)
  train_scores <- all_g_scores[all_g_scores$split_type == "train", ]
  avg_train_scores <- aggregate(g ~ subject_id, train_scores, mean)
  avg_train_scores$split_type <- "train_averaged"
  
  # For timepoint IDs in averaged scores, we'll need to handle this differently
  # since each subject has multiple train timepoints
  
  return(list(
    all_scores = all_g_scores,
    test_only = test_scores_only,
    train_averaged = avg_train_scores
  ))
}

# Extract the scores
nested_cv_scores <- combine_nested_cv_scores()

# Save different versions
write.csv(nested_cv_scores$all_scores, 
          file = file.path(save_path, "all_g_factor_scores_nested_cv_complete.csv"), 
          row.names = FALSE)

write.csv(nested_cv_scores$test_only, 
          file = file.path(save_path, "all_g_factor_scores_nested_cv_test_only.csv"), 
          row.names = FALSE)

write.csv(nested_cv_scores$train_averaged, 
          file = file.path(save_path, "all_g_factor_scores_nested_cv_train_averaged.csv"), 
          row.names = FALSE)
```

# Analyze longitudinal patterns

```{r longitudinal, fig.width=10, fig.height=6}
# Updated longitudinal analysis for nested CV
analyze_nested_cv_longitudinal <- function() {
  # Load the processed data with all information
  all_data_file <- file.path(save_path, "all_longitudinal_cognitive_data_processed.csv")
  if (!file.exists(all_data_file)) {
    cat("Processed data file not found. Skipping longitudinal analysis.\n")
    return(NULL)
  }
  
  all_data <- read.csv(all_data_file)
  
  # Option 1: Use only test scores (each subject appears once)
  cat("\n=== LONGITUDINAL ANALYSIS USING TEST SCORES ONLY ===\n")
  test_merged <- merge(nested_cv_scores$test_only, all_data, 
                      by.x = "eid", by.y = "timepoint_id", all.x = TRUE)
  
  # Plot using test scores only
  p1 <- ggplot(test_merged, aes(x = age_years, y = g)) +
    geom_point(size = 2, alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "blue") +
    labs(title = "G-Factor Scores by Age (Test Scores Only - Nested CV)",
         subtitle = "Each subject appears exactly once",
         x = "Age (years)",
         y = "G-Factor Score") +
    theme_minimal()
  
  print(p1)
  
  # Option 2: Use all scores but account for repeated measures
  cat("\n=== LONGITUDINAL ANALYSIS USING ALL SCORES ===\n")
  all_merged <- merge(nested_cv_scores$all_scores, all_data, 
                     by.x = "eid", by.y = "timepoint_id", all.x = TRUE)
  
  # For subjects with multiple longitudinal timepoints, show trajectories
  # But note that some "trajectories" will be artifacts of the CV structure
  subjects_with_multiple_real_timepoints <- names(table(all_data$subject_id))[table(all_data$subject_id) > 1]
  
  if (length(subjects_with_multiple_real_timepoints) > 0) {
    multi_timepoint_data <- all_merged[all_merged$subject_id %in% subjects_with_multiple_real_timepoints,]
    
    # Separate by split type for clearer visualization
    p2 <- ggplot(multi_timepoint_data, aes(x = age_years, y = g, group = subject_id)) +
      geom_line(alpha = 0.3) +
      geom_point(aes(color = split_type, shape = factor(outer_fold)), size = 2) +
      labs(title = "G-Factor Trajectories (Subjects with Multiple Real Timepoints)",
           subtitle = "Colors indicate train/test split, shapes indicate outer fold",
           x = "Age (years)",
           y = "G-Factor Score",
           color = "Split Type",
           shape = "Outer Fold") +
      theme_minimal()
    
    print(p2)
  }
  
  # Option 3: Model validation - compare train vs test performance
  cat("\n=== MODEL VALIDATION ANALYSIS ===\n")
  
  # For each outer fold, compare train and test score distributions
  validation_stats <- data.frame()
  
  for (i in 0:4) {
    fold_train <- nested_cv_scores$all_scores[nested_cv_scores$all_scores$outer_fold == i & 
                                              nested_cv_scores$all_scores$split_type == "train", ]
    fold_test <- nested_cv_scores$all_scores[nested_cv_scores$all_scores$outer_fold == i & 
                                             nested_cv_scores$all_scores$split_type == "test", ]
    
    if (nrow(fold_train) > 0 && nrow(fold_test) > 0) {
      validation_stats <- rbind(validation_stats, data.frame(
        outer_fold = i,
        train_mean = mean(fold_train$g, na.rm = TRUE),
        train_sd = sd(fold_train$g, na.rm = TRUE),
        test_mean = mean(fold_test$g, na.rm = TRUE),
        test_sd = sd(fold_test$g, na.rm = TRUE),
        train_n = nrow(fold_train),
        test_n = nrow(fold_test)
      ))
    }
  }
  
  # Plot validation statistics
  if (nrow(validation_stats) > 0) {
    library(reshape2)
    means_long <- melt(validation_stats[, c("outer_fold", "train_mean", "test_mean")], 
                      id.vars = "outer_fold", variable.name = "split", value.name = "mean_g")
    
    p3 <- ggplot(means_long, aes(x = factor(outer_fold), y = mean_g, fill = split)) +
      geom_bar(stat = "identity", position = "dodge") +
      labs(title = "Mean G-Factor Scores by Fold and Split Type",
           x = "Outer Fold",
           y = "Mean G-Factor Score",
           fill = "Split Type") +
      theme_minimal()
    
    print(p3)
    
    # Print validation statistics
    print(validation_stats)
  }
  
  # Save merged datasets
  write.csv(test_merged, 
            file = file.path(save_path, "longitudinal_g_factor_nested_cv_test_only.csv"), 
            row.names = FALSE)
  
  write.csv(all_merged, 
            file = file.path(save_path, "longitudinal_g_factor_nested_cv_complete.csv"), 
            row.names = FALSE)
  
  return(list(test_merged = test_merged, all_merged = all_merged, validation_stats = validation_stats))
}

# Run the analysis
longitudinal_results <- analyze_nested_cv_longitudinal()
```

# Conclusion

This analysis implements a regularized (small constant to covar matrix) g-factor model which treats all variables as continuous while providing improved numerical stability.
